\chapter{Regulatory Agents}
\label{ch:reg-agents}

\section{Why the regulations agents live in \texttt{src/intelligence}}

The repository does not contain a standalone
\filepath{src/regulations/} service. The concrete runtime surface for
regulations obligations lives in and around:
\begin{itemize}
  \item \filepath{src/intelligence/ai-core-pal/}
  \item \filepath{src/intelligence/monitoring/}
  \item \filepath{src/intelligence/analytics/}
\end{itemize}

The regulations development and runtime-monitor rule packs therefore
belong under \filepath{src/intelligence/}.

\section{Required regulatory development agent}

The development pack at \filepath{src/intelligence/.clinerules} must
govern any work that claims to implement the regulations domain. It
must be schema-backed by:
\begin{itemize}
  \item \filepath{docs/schema/regulations/requirement.schema.json}
  \item \filepath{docs/schema/regulations/conformance-tool.schema.json}
  \item \filepath{docs/schema/regulations/agent-identity.schema.json}
  \item \filepath{docs/schema/regulations/request-identity.schema.json}
  \item \filepath{docs/schema/regulations/audit-event.schema.json}
  \item \filepath{docs/schema/regulations/capability-monitoring-metrics.schema.json}
\end{itemize}

It must point agents to the live implementation surface, including:
\begin{itemize}
  \item \filepath{src/intelligence/ai-core-pal/mcp_server/btp_pal_mcp_server.py}
  \item \filepath{src/intelligence/ai-core-pal/data_products/registry.yaml}
  \item \filepath{src/intelligence/analytics/anomaly_detector.py}
  \item \filepath{src/intelligence/monitoring/kill_switch_monitor.py}
  \item \filepath{src/intelligence/ai-core-pal/tests/test_mcp_server.py}
\end{itemize}

\section{Required regulatory runtime-monitor agent}

The runtime-monitor pack at
\filepath{src/intelligence/.clinerules.runtime-monitor} must supervise:
\begin{itemize}
  \item allow-list bypass and undeclared tool use;
  \item missing or partial identity envelopes;
  \item audit reservation failures and orphaned completions;
  \item monitoring metric gaps;
  \item schema-violation surges;
  \item refusal-rate spikes, latency spikes, topic drift, and related
    degradation triggers;
  \item circuit-break conditions and alert routing.
\end{itemize}

\section{Minimum acceptance criteria for the regulations agent pair}

\begin{enumerate}
  \item The development pack cites the regulations specification and
    schema files as source of truth.
  \item The runtime-monitor pack cites the monitoring and identity
    chapters as source of truth.
  \item Both packs name concrete code paths and test entry points that
    exist in the repository.
  \item A validation harness checks both packs for structural
    completeness and benchmark quality.
  \item The combined pair benchmarks at least as strongly as the
    existing \filepath{src/generativeUI/.clinerules} baseline on
    operational specificity.
\end{enumerate}

\section{Ingress overlay requirements}
\label{sec:ingress-overlay}

The primary regulatory pair lives under \filepath{src/intelligence/},
because that is where the core governance runtime lives. However, if a
service in \filepath{src/generativeUI/gateway/} is the primary entry
point for external traffic \textbf{AND} it handles
authentication/authorization token extraction, then a companion pack at
\filepath{src/generativeUI/gateway/.clinerules.runtime-monitor} is
\textbf{MANDATORY}, not merely recommended. This mandatory pack SHOULD be
automatically generated from the primary intelligence pack plus a
gateway-specific overlay template to minimize maintenance overhead.

\subsection{Mandatory ingress enforcement}

When the gateway is the external traffic entry point, the ingress
overlay MUST enforce:

\begin{enumerate}
  \item \textbf{Header propagation validation}: The monitor MUST verify
    that required headers (\texttt{X-Request-ID}, \texttt{X-Agent-Identity},
    \texttt{X-Correlation-ID}) are present and correctly formatted before
    requests reach internal services.
  \item \textbf{Early rejection of invalid requests}: Requests missing
    required identity claims MUST be rejected at the gateway level with
    appropriate HTTP status codes (401/403) and audit logging.
  \item \textbf{Identity envelope integrity}: The gateway MUST validate
    that identity tokens have not been tampered with and originate from
    trusted issuers.
  \item \textbf{Rate limiting enforcement}: Per-identity rate limits
    MUST be enforced at the gateway to prevent resource exhaustion attacks.
\end{enumerate}

\subsection{Ingress overlay minimum contents}

The ingress runtime-monitor pack at
\filepath{src/generativeUI/gateway/.clinerules.runtime-monitor} MUST
include:

\begin{table}[htbp]
\centering
\footnotesize
\caption{Ingress overlay minimum requirements.}
\label{tab:ingress-requirements}
\begin{tabularx}{\textwidth}{@{}l X@{}}
\toprule
\textbf{Section} & \textbf{Required content} \\
\midrule
Evidence sources & Gateway access logs, authentication service logs, rate-limiter metrics. \\
Critical conditions & Missing identity headers, invalid tokens, rate limit exceeded, upstream timeout. \\
Alert routing & Security team for identity violations; platform team for availability issues. \\
Circuit-break triggers & 3+ consecutive authentication failures from same source in 1 minute. \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Inheritance from primary regulatory pair}

The ingress overlay MUST inherit from (not replace) the primary
regulatory runtime-monitor at
\filepath{src/intelligence/.clinerules.runtime-monitor}. Specifically:
\begin{itemize}
  \item All safety-critical flags from the parent MUST be preserved or
    tightened (see \cref{sec:conflict-detection}).
  \item The ingress overlay MAY add gateway-specific conditions but
    MUST NOT remove or weaken inherited conditions.
  \item Audit events from the gateway MUST conform to the same
    \filepath{docs/schema/regulations/audit-event.schema.json} as the
    core governance runtime.
\end{itemize}

\subsection{Exception for non-external gateways}

If the gateway at \filepath{src/generativeUI/gateway/} handles
\emph{only} internal service-to-service traffic (not external user
requests), the ingress overlay MAY be designated as \texttt{OPTIONAL}
with documented justification in the rule pack frontmatter:
\begin{verbatim}
# Ingress overlay: OPTIONAL
# Justification: Gateway handles internal traffic only.
# External entry point: src/external-gateway/
# Reviewed by: [Approver ID]
# Review date: [Date]
\end{verbatim}
