% !TEX program = xelatex
\documentclass[11pt,a4paper,twoside]{report}

\usepackage{xcolor}
\usepackage[a4paper,margin=28mm]{geometry}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{listings}
\usepackage[hyphens]{url}
\usepackage{tcolorbox}
\usepackage{tikz}
\usepackage[table]{colortbl}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc,chains,
decorations.pathmorphing,backgrounds}

\newtcolorbox{adrbox}[2][]{colback=blue!5,colframe=blue!75!black,fonttitle=\bfseries,title=ADR: #2,#1}

% --- Semantic Hyperlinking --------------------------------------------------
\usepackage{xr-hyper}
\usepackage[hidelinks]{hyperref}

% External references to peer specifications
\externaldocument[reg:]{../regulations/regulations-spec}
\externaldocument[tb:]{../tb/tb-review-spec}
\externaldocument[sim:]{../simula/simula-training-spec}
\externaldocument[ar:]{../arabic/arabic-ap-spec}

% --- Fonts -------------------------------------------------------------------
\usepackage{fontspec}
\setmainfont{Arial}
\usepackage[capitalize,noabbrev]{cleveref}

\definecolor{codebg}{gray}{0.96}
\definecolor{sapblue}{HTML}{0B5FA5}
\definecolor{sapgreen}{HTML}{00695C}
\definecolor{sapamber}{HTML}{B8360F}
\definecolor{sapgrey}{HTML}{546E7A}
\definecolor{sapmidnight}{HTML}{1C2B39}
\definecolor{sapgold}{HTML}{F0AB00}

\lstdefinelanguage{jsonx}{
basicstyle=\ttfamily\small,
columns=fullflexible,
breaklines=true,
showstringspaces=false,
comment=[l]{//},
morestring=[b]",
stringstyle=\color{sapgreen},
keywordstyle=\color{sapblue},
}

\lstdefinelanguage{yamlx}{
keywords={true,false,null,y,n},
keywordstyle=\color{sapblue}\bfseries,
basicstyle=\ttfamily\small,
sensitive=false,
comment=[l]{\#},
commentstyle=\color{sapgreen}\ttfamily,
stringstyle=\color{sapamber}\ttfamily,
moredelim=[l][\color{sapmidnight}]{\:},
morestring=[b]',
morestring=[b]",
}

\onehalfspacing
\setlist{topsep=2pt,itemsep=1pt,parsep=0pt}
\setlength{\emergencystretch}{3em}
\tolerance=1000
\hbadness=10000

% --- Global Command Overrides -----------------------------------------------
\newcommand{\corpus}{}
\newcommand{\tool}{}
\newcommand{\stateid}[1]{\textsc{\lowercase{#1}}}
\newcommand{\ar}[1]{#1}

% --- Glossary Helper Commands -----------------------------------------------
\definecolor{pathcolor}{HTML}{004D40}
\newcommand{\schemapath}[1]{\texttt{\textcolor{pathcolor}{#1}}}
\newcommand{\enumval}[1]{\texttt{\textcolor{sapgreen}{#1}}}
\newcommand{\fieldref}[1]{\texttt{\textcolor{sapblue}{#1}}}

\begin{document}

\title{\Huge Arabic AP Domain \\
\textbf{Business Requirements \& Lineage} \\
\large HITL Traceability Specification \\
\small Ref: \texttt{ARA-REQ-LIN-2026-v1.0}}
\author{SAP-OSS Business Architecture Team}
\date{April 2026}

\maketitle

% --- Incorporated Executive Summary ---

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

This specification defines the human-in-the-loop (HITL) traceability for the Arabic AP domain. It provides a formal mapping between raw regulatory source documents, derived business requirements, and the technical implementation specifications.

The document is designed for business stakeholders, auditors, and legal reviewers to verify that the AI-augmented pipeline respects the underlying mandates of ZATCA (Saudi), NBR (Bahrain), and other regional authorities.
\chapter{Source Material Registry}
\label{ch:source-registry}

This chapter catalogues the primary source documents used to derive the business requirements for the Arabic AP domain.

\begin{longtable}{p{5cm} p{3cm} p{6cm}}
\caption{Source Document Inventory} \\
\toprule
\textbf{File Name} & \textbf{Regulatory Body} & \textbf{Primary Concern} \\
\midrule
\endfirsthead
\toprule
\textbf{File Name} & \textbf{Regulatory Body} & \textbf{Primary Concern} \\
\midrule
\endhead
\texttt{Accounts Payable DOI.docx} & SCB Bahrain / NBR & Delegation of Invoicing (DOI) \& Controls \\
\texttt{Saudi - OTP Entry processing...} & ZATCA (Saudi) & VAT Q3'25 processing rules \\
\texttt{EDAA Annual Fee Invoice...} & EDAA & Arabic/English Invoice layout (Standard) \\
\texttt{Arabic-Simplified-Invoice.pdf} & Local Authority & Arabic-only simplified invoice structure \\
\bottomrule
\end{longtable}
\chapter{Transformation Logic}
\label{ch:transformation}

\section{Compliance Lifecycle: Source to Evidence}
The following diagram visualizes the value stream for requirements in the SAP OSS platform. It ensures that every line of code is anchored in a verified business mandate.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
node distance=1.5cm,
auto,
block/.style={rectangle, draw, rounded corners, fill=sapmidnight!5, text width=8em, text centered, minimum height=3em, font=\scriptsize},
arrow/.style={->, >=latex, thick, sapmidnight}
]
\node (source) [block] {\textbf{Source Material}\\(PDF/DOCX/XLS)};
\node (analyst) [block, right of=source, xshift=2.5cm] {\textbf{GFS Analyst}\\(Heuristic Extraction)};
\node (req) [block, right of=analyst, xshift=2.5cm] {\textbf{Business Req}\\(BR Catalogue)};
\node (spec) [block, below of=req, yshift=-1cm] {\textbf{Implementation Spec}\\(Ch. 12 Tasks)};
\node (test) [block, left of=spec, xshift=-2.5cm] {\textbf{Test Evidence}\\(Verification Blocks)};
\node (signoff) [block, left of=test, xshift=-2.5cm] {\textbf{HITL Sign-off}\\(Final Acceptance)};

\draw [arrow] (source) -- (analyst);
\draw [arrow] (analyst) -- (req);
\draw [arrow] (req) -- (spec);
\draw [arrow] (spec) -- (test);
\draw [arrow] (test) -- (signoff);
\end{tikzpicture}
\caption{The end-to-end compliance value stream.}
\end{figure}

\section{From PDF to Machine-Readable (YAML/JSON)}
The transformation from narrative prose to technical enforcement uses a semi-automated "Ground Truth" process. \cref{tab:artifact-linkage} maps the raw sources to their corresponding machine-readable counterparts.

\begin{table}[htbp]
\centering
\footnotesize
\caption{Source Material to Machine-Readable Linkage}
\label{tab:artifact-linkage}
\begin{tabularx}{\textwidth}{@{}l X X@{}}
\toprule
\textbf{Source Material (Raw)} & \textbf{Machine-Readable Artifact} & \textbf{Registry Location} \\
\midrule
\texttt{Accounts Payable DOI.docx} & \texttt{bh-ap-doi.json} & \texttt{structured/policies/} \\
\texttt{Saudi - OTP Entry...} & \texttt{ksa-vat-q3.yaml} & \texttt{structured/rules/} \\
\texttt{Specification Addendum} & \texttt{spec-addendum.yaml} & \texttt{arabic/machine-readable/} \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Extraction Methodology}
\begin{enumerate}
\item \textbf{Heuristic Extraction:} Initial requirement identification by GFS Analysts based on domain knowledge.
\item \textbf{Schema Normalization:} Mapping requirements to the \texttt{invoice.schema.json} and \texttt{vat-checklist.schema.json} defined in the technical spec.
\item \textbf{Formalization:} Converting narrative rules (e.g., "Full tax invoice must contain TIN") into JSON-Logic predicates for automated enforcement.
\end{enumerate}
\chapter{Business Requirements Catalogue}
\label{ch:requirements}

\begin{longtable}{p{2cm} p{5cm} p{7cm}}
\caption{Consolidated Business Requirements} \\
\toprule
\textbf{Req ID} & \textbf{Business Rule} & \textbf{Source Document Reference} \\
\midrule
\endfirsthead
\toprule
\textbf{Req ID} & \textbf{Business Rule} & \textbf{Source Document Reference} \\
\midrule
\endhead
\rowcolor{sapblue!5}
\textbf{BR-OCR-01} & Character precision for currency must exceed 99.5\%. & \texttt{Saudi - OTP Entry processing...} (p. 2) \\
\textbf{BR-VAT-SA} & Simplified vs Full Invoice determination based on 1000 SAR threshold. & ZATCA VAT Guidelines (Implicit in Saudi OTP) \\
\rowcolor{sapblue!5}
\textbf{BR-DOI-07} & Evidence-based seven-control sign-off for Bahrain. & \texttt{Accounts Payable DOI.docx} (Section 4) \\
\textbf{BR-TRN-01} & Translation must maintain semantic mapping for legal entity names. & \texttt{EDAA Annual Fee Invoice...} \\
\bottomrule
\end{longtable}
\chapter{Requirement-to-Specification Lineage}
\label{ch:lineage}

\begin{longtable}{p{2cm} p{6cm} p{6cm}}
\caption{Technical Traceability Matrix} \\
\toprule
\textbf{Req ID} & \textbf{Technical Implementation} & \textbf{Implementation Spec Ref} \\
\midrule
\rowcolor{sapamber!5}
\textbf{BR-OCR-01} & Confidence threshold calibration pilot. & \texttt{arabic-ap-spec} Ch. 10 \\
\textbf{BR-VAT-SA} & Rule ID \texttt{VAT-SA-SI-001} in VAT Engine. & \texttt{arabic-ap-spec} Ch. 8 \\
\rowcolor{sapamber!5}
\textbf{BR-DOI-07} & 28-state workflow wiring in \texttt{ap.workflow.advance}. & \texttt{arabic-ap-spec} Ch. 5 \& 6 \\
\textbf{BR-TRN-01} & Arabic--English translation gate calibration. & \texttt{arabic-ap-spec} Ch. 3 \\
\bottomrule
\end{longtable}

% --- NEW Detailed Analysis Chapters ---
\chapter{Source Analysis: Bahrain AP DOI}
\label{ch:analysis-bh-doi}

\section{Document Context}
The \texttt{Accounts Payable DOI.docx} (Bahrain) defines the delegation of authority and mandatory controls for invoice processing. This document was processed into \texttt{structured/policies/bh-ap-doi.json}.

\section{Detailed Requirement Mapping}

\begin{longtable}{p{5cm} p{5cm} p{5cm}}
\caption{Bahrain DOI: Source-to-Requirement Lineage} \\
\toprule
\textbf{Source Paragraph} & \textbf{Business Requirement} & \textbf{Formal Requirement (Spec)} \\
\midrule
\endfirsthead
\toprule
\textbf{Source Paragraph} & \textbf{Business Requirement} & \textbf{Formal Requirement (Spec)} \\
\midrule
\endhead
Section 4.1: ``All invoices must be matched against a valid Purchase Order prior to processing.'' & \textbf{BR-DOI-M1:} Mandatory PO-Matching for standard AP flow. & \texttt{arabic-ap-spec} Ch. 5 (PO-EPROC state machine). \\
\midrule
Section 4.2: ``Any variance > 5\% between PO and Invoice must be routed to the Finance Head.'' & \textbf{BR-DOI-V1:} Automated variance threshold detection. & \texttt{arabic-ap-spec} Ch. 6 (Variance Control C3). \\
\midrule
Section 4.7: ``The final payment file must be hashed and signed by two distinct personas.'' & \textbf{BR-DOI-S1:} Dual-Persona Cryptographic Sign-off. & \texttt{arabic-ap-spec} Ch. 12 (HITL Gate: Payment Instruction). \\
\bottomrule
\end{longtable}

\section{Machine-Readable Evidence (Fragment)}
The following YAML fragment was extracted from the DOI to drive the 28-state workflow engine:

\begin{lstlisting}[language=jsonx]
{
"policy_id": "BH-DOI-2026",
"controls": [
{
"id": "C3",
"description": "PO-Invoice Variance check",
"threshold": 0.05,
"escalation_persona": "finance_head"
}
]
}
\end{lstlisting}
\chapter{Source Analysis: Saudi VAT (ZATCA)}
\label{ch:analysis-sa-vat}

\section{Document Context}
The \texttt{Saudi - OTP Entry processing...} (ZATCA) defines the rules for electronic invoicing (E-Invoicing) and simplified vs. full tax invoice criteria. This was processed into \texttt{structured/rules/ksa-vat-q3.yaml}.

\section{Detailed Requirement Mapping}

\begin{longtable}{p{5cm} p{5cm} p{5cm}}
\caption{Saudi ZATCA: Source-to-Requirement Lineage} \\
\toprule
\textbf{Source Provision} & \textbf{Business Requirement} & \textbf{Technical Rule (Spec Ch. 8)} \\
\midrule
\endfirsthead
\toprule
\textbf{Source Provision} & \textbf{Business Requirement} & \textbf{Technical Rule (Spec Ch. 8)} \\
\midrule
\endhead
Article 53(b): ``Simplified Tax Invoices are issued for transactions < 1,000 SAR.'' & \textbf{BR-VAT-SI:} Determination of Simplified vs Full invoice based on threshold. & Rule \texttt{VAT-SA-THRESHOLD}: \texttt{gross\_amount < 1000} \\
\midrule
Article 54: ``Full Tax Invoices must display the Buyer's VAT Registration Number.'' & \textbf{BR-VAT-FULL:} Mandatory TRN/TIN field extraction for Full invoices. & Rule \texttt{VAT-SA-MANDATORY}: \texttt{exists(buyer.trn)} \\
\midrule
Phase 2 (ZATCA Integration): ``Invoice must contain a Phase-2 QR code including the ECDSA signature.'' & \textbf{BR-VAT-QR:} Extraction and verification of ZATCA QR signatures. & \texttt{arabic-ap-spec} Ch. 10 (OCR Pilot: QR calibration). \\
\bottomrule
\end{longtable}

\section{Machine-Readable Evidence (Fragment)}
The following YAML extraction defines the logic evaluated by the VAT engine for Saudi Arabian invoices:

\begin{lstlisting}[language=yamlx]
domain: "ksa-vat"
rules:
- id: "THRESHOLD_CHECK"
type: "classification"
predicate: "invoice.total_gross < 1000"
target_schema: "simplified_invoice"
- id: "TRN_CHECK"
type: "validation"
requirement: "buyer.trn must match ^3[0-9]{14}$"
\end{lstlisting}

% --- Incorporated Universal Chapters ---
\chapter{Global Wave-Dependency Manifest}
\label{ch:global-dependencies}

\section{Critical Path Visualization (Waves 1--3)}
The following chart defines the normative sequencing for implementation agents. A task in a later wave cannot be marked \textsc{complete} until its prerequisites in an earlier wave are validated.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
node distance=1.5cm,
auto,
wavebox/.style={rectangle, draw, rounded corners, fill=sapmidnight!10, text width=10em, text centered, minimum height=3em, font=\scriptsize\bfseries},
dep/.style={->, >=latex, thick, sapblue},
critical/.style={->, >=latex, very thick, sapamber}
]
% --- Wave 1: Foundations ---
\node (r1) [wavebox, fill=sapgrey!20] {\textcolor{sapmidnight}{REG-W1: Identity \& Audit Schema}};
\node (s1) [wavebox, below of=r1, yshift=-0.5cm] {SIM-W1: HANA Extraction Pipeline};
\node (a1) [wavebox, below of=s1, yshift=-0.5cm] {ARA-W1: Invoice Intake \& OCR};
\node (t1) [wavebox, below of=a1, yshift=-0.5cm] {TB-W1: Identity Wiring};

% --- Wave 2: Core Build-out ---
\node (r2) [wavebox, right of=r1, xshift=4.5cm, fill=sapmidnight!20] {\textcolor{sapmidnight}{REG-W2: Governance Runtime}};
\node (s2) [wavebox, right of=s1, xshift=4.5cm] {SIM-W2: Taxonomy \& Synthesis};
\node (a2) [wavebox, right of=a1, xshift=4.5cm] {ARA-W2: Multi-Country VAT};
\node (t2) [wavebox, right of=t1, xshift=4.5cm] {TB-W2: Controls \& Workflow};

% --- Wave 3: Hardening ---
\node (r3) [wavebox, right of=r2, xshift=4.5cm, fill=sapmidnight!30] {\textcolor{sapmidnight}{REG-W3: Conformance Suite}};
\node (s3) [wavebox, right of=s2, xshift=4.5cm] {SIM-W3: 85\% Coverage Test};
\node (a3) [wavebox, right of=a2, xshift=4.5cm] {ARA-W3: Calibration Pilot};
\node (t3) [wavebox, right of=t2, xshift=4.5cm] {TB-W3: HITL Dashboard};

% --- Cross-Domain Dependencies (The Critical Path) ---
\draw [critical] (r1) -- node [above, font=\tiny, align=center] {Identity Envelope\\Invariants} (a1.north east);
\draw [critical] (r1) -- (t1.north east);
\draw [dep] (s1.east) -- node [above, font=\tiny, xshift=-0.5cm] {Training Fixtures} (t2.west);
\draw [dep] (s1.east) -- (a2.west);
\draw [dep] (r2.south) -- node [left, font=\tiny, yshift=0.3cm] {Policy Gate} (a2.north);
\draw [dep] (t2.east) -- node [above, font=\tiny] {Audit Evidence} (r3.west);

% --- Wave Transitions ---
\draw [->, dashed, sapgrey] (r1.east) -- (r2.west);
\draw [->, dashed, sapgrey] (r2.east) -- (r3.west);

% --- Legend ---
\node[draw, sapamber, very thick, label=right:\scriptsize Critical Prerequisite] at (0,-7.5) {};
\node[draw, sapblue, thick, label=right:\scriptsize Data Dependency, xshift=4cm] at (0,-7.5) {};

\end{tikzpicture}
\caption{Cross-specification dependency graph: The blueprint for platform rollout.}
\end{figure}

\section{Inter-Domain Prerequisite Matrix}

\begin{table}[htbp]
\centering
\footnotesize
\caption{Normative Cross-Domain Prerequisites.}
\label{tab:prerequisites}
\begin{tabularx}{\textwidth}{@{}l l l X@{}}
\toprule
\textbf{Target Task} & \textbf{Prerequisite} & \textbf{Ref ID} & \textbf{Rationale} \\
\midrule
\rowcolor{sapmidnight!5}
\textbf{ARA-W1} & REG-W1 (Identity) & \texttt{REG-IDENTITY-01} & Invoice intake must be audit-traced from Day 1. \\
\textbf{TB-W2}  & SIM-W1 (Extraction) & \texttt{SIM-EXTRACT-01} & Anomaly detection requires HANA schema metadata. \\
\rowcolor{sapmidnight!5}
\textbf{REG-W2} & SIM-W1 (Registry) & \texttt{SIM-REGISTRY-01} & Governance monitoring relies on the schema registry. \\
\textbf{ARA-W2} & REG-W2 (Pillar 1) & \texttt{REG-MGF-P1} & VAT release is an HITL gate governed by MAS Pillar 1. \\
\rowcolor{sapmidnight!5}
\textbf{TB-W3}  & ARA-W2 (Posted Invoices) & \texttt{ARA-AP-POST} & Review dashboard depends on posted AP data from Arabic domain. \\
\bottomrule
\end{tabularx}
\end{table}

\begin{adrbox}{Dependency Enforcement}
Implementation agents are strictly forbidden from "Fast-Forwarding" into Wave 2 if Wave 1 critical prerequisites (marked in \textcolor{sapamber}{\textbf{Gold}}) are in a \textsc{gap} or \textsc{partial} state.
\end{adrbox}



\end{document}
