\chapter{Agent Implementation Instructions}
\label{ch:agent-impl}

\section{Purpose and audience}
This chapter is the self-contained, intent-based instruction set for
implementing the \emph{Simula training data framework} domain inside
the unified SAP OSS platform. It is written for LLM-driven and human
implementation agents who will extend existing services to realise
the specification. It assumes only this PDF; every platform
decision, integration contract, and governance obligation the Simula
workstream depends on is restated here so that no companion document
is required. Existing sections~\ref{chap:overview}--\ref{sec:sim-roadmap}
remain the normative source for the Simula domain itself; this
chapter tells future agents \emph{how} to build what those sections
describe, \emph{where} to put it in the codebase, and \emph{how} it
plugs into the wider platform.

\input{../shared-intent-core.tex}

\section{FMEA: Failure Mode and Effects Analysis for AI}
\label{sec:fmea-ai}

The following analysis identifies critical AI failure modes within the Simula pipeline and their deterministic mitigations.

\begin{table}[htbp]
\centering
\footnotesize
\caption{FMEA for Simula AI Pipeline.}
\label{tab:simula-fmea}
\begin{tabularx}{\textwidth}{@{}l X X X@{}}
\toprule
\textbf{Failure Mode} & \textbf{Symptom} & \textbf{Impact} & \textbf{Technical Mitigation} \\
\midrule
Critic Collusion & Both critics pass a low-quality SQL query. & High (Model Bias) & Inject a 'Gold Standard' probe query every 100 samples. \\
Schema Contraction & LLM ignores complex JOIN paths in the taxonomy. & Medium (Coverage) & Taxonomic coverage check against Graph-DB baseline. \\
Elo-Score Drift & Calibrated complexity scores diverge from human ratings. & Medium (UI/UX) & Monthly Elo-recalibration against human-label set. \\
\bottomrule
\end{tabularx}
\end{table}

\section{This domain in context}
\label{sec:agent-domain-context}
Simula is the training-data workstream among the four domains. Its
primary persona is the Training Engineer, whose working view is the
Simula console surfaced through the shared workspace shell and gated
by Simula entitlements. The domain's entry points are the schema
sets and fixtures exported by Arabic AP, TB, and regulations, plus
the HANA Cloud catalogue read via the extraction pipeline; its exit
points are \texttt{TrainingExample} records emitted to the training
corpus and Elo-calibrated complexity reports. Every artefact is
validated against the Simula-owned schemas in
\texttt{structured/schema/} before it leaves the pipeline. The
domain is read-only with respect to the other three domains at the
schema layer: it never mutates their records, and it never calls
their runtime MCP tools. The horizontal regulations wrapper governs
Simula calls like any other domain.

\section{Schema contribution}
\label{sec:agent-schema-contrib}
Simula owns the schemas catalogued in
chapter~\ref{chap:schema}, including the training-example record,
the schema-registry record, the taxonomy structure, and the
enumerations listed there. These schemas live under
\texttt{docs/latex/specs/simula/} as normative text and under
\texttt{structured/schema/} as the JSON Schema artefacts consumed by
the pipeline at \texttt{src/training/pipeline/}. Registration in the
unified registry is additive: each schema is listed in
\texttt{src/intelligence/ai-core-pal/data\_products/registry.yaml}
with its URI, owning domain (\texttt{simula}), security class
(\texttt{confidential}), and LLM routing (\texttt{vllm-only}),
following the ODPS~4.1 pattern already documented for HANA training
data products in that file. Simula also consumes the schemas owned
by Arabic, TB, and regulations as inputs; the contract is strictly
schema-in, \texttt{TrainingExample}-out. Simula performs no schema
merges; the two \texttt{requirement.schema.json} files in TB and
regulations remain distinct when Simula reads them.

\section{Persona and MCP allow-list}
\label{sec:agent-persona}
\begin{itemize}
  \item \textbf{Persona id:} \texttt{training-engineer}.
  \item \textbf{Roles:} schema extraction, taxonomy authoring,
    training-data synthesis supervision, critic-calibration review,
    corpus export.
  \item \textbf{Entitlements:} read/write on the training-example
    record, schema-registry record, and taxonomy structure; read on
    every other domain's schemas (Arabic, TB, regulations);
    run-execute on Simula pipeline stages described in
    chapter~\ref{chap:generator}.
  \item \textbf{UI route:} \texttt{/simula/\{tenant\}/console} on the
    shared workspace shell; surfaced only to callers whose
    entitlement set includes \texttt{training-engineer}.
  \item \textbf{MCP tool allow-list:}
    \texttt{simula.schema.extract}, \texttt{simula.taxonomy.build},
    \texttt{simula.generate.examples},
    \texttt{simula.critic.calibrate},
    \texttt{simula.evaluate.coverage},
    \texttt{simula.evaluate.diversity},
    \texttt{simula.corpus.export}. All other MCP tools are denied by
    default per MAS MGF pillar~1.
\end{itemize}

\section{Implementation tasks for this domain}
\label{sec:agent-tasks}
The tasks below are intent-based: each task states the outcome a
future agent must achieve and the location in the existing codebase
where the work belongs. None of them contains code; each task binds
to the Simula sections that define the required behaviour.

\subsubsection{Task S1 --- HANA schema extraction.}
\textbf{Intent.} Realise the HANA schema extraction pipeline in
chapter~\ref{chap:extraction}, covering MCP discovery, table
enumeration, column extraction, and registry population.

\begin{adrbox}{Why Direct SQL Metadata?}
We chose direct SQL-metadata extraction over high-level MCP discovery for the baseline to ensure we capture SAP-specific table types (Global Temporary, Row-Store) that standard MCP tools often omit.
\end{adrbox}

\textbf{In-scope.} Each of the four stages in the pipeline figure
(\ref{fig:extraction-pipeline}); schema registry writer; per-table
dimension/measure classification.
\textbf{Out-of-scope.} Non-HANA backends.
\textbf{Paths to extend.} \texttt{src/training/pipeline/} extraction
module; registry writer that emits entries conforming to the
schema-registry record.
\textbf{Definition of done.} Running extraction against a HANA
Cloud catalogue produces a registry populated with every in-scope
table; registry entries validate against the schema-registry record.

\textbf{Autonomous Verification (Agent Logic):}
\begin{lstlisting}[language=jsonx]
// Verify Task S1
{
  "test_fixture": "tests/fixtures/simula/hana_mock.sql",
  "expected_outcome": "table_count == 14",
  "validation_command": "python src/simula/extract.py --verify-registry"
}
\end{lstlisting}

\textbf{Verification.} Pipeline dry-run against the fixture store
described in chapter~\ref{ch:test-fixtures} produces the expected
registry entries.

\subsubsection{Task S2 --- Taxonomy generation (Algorithm 1).}
\textbf{Intent.} Implement Algorithm~1 (reasoning-driven taxonomy
generation) from chapter~\ref{chap:taxonomy} in full, including all
four steps (factor identification, breadth-first expansion,
best-of-N proposal, critic refinement).
\textbf{In-scope.} Per-step module with the checkpoints
in~\ref{fig:taxonomy-stages}; taxonomy structure writer.
\textbf{Out-of-scope.} Manual taxonomy authoring UI.
\textbf{Paths to extend.} Taxonomy engine colocated with the
pipeline.
\textbf{Definition of done.} Running Algorithm~1 against an extracted
registry produces a taxonomy that validates against the taxonomy
structure schema and passes the coverage check in
chapter~\ref{chap:evaluation}.
\textbf{Verification.} Taxonomy engine unit tests pass; the
coverage evaluator reports coverage within the chapter's target
band.

\subsubsection{Task S3 --- Agentic data synthesis (Algorithm 2).}
\textbf{Intent.} Implement Algorithm~2 (agentic data synthesis) from
chapter~\ref{chap:generator}, including all four steps (mix
sampling, meta-prompt generation, complexification, example
generation) and the double-critic validation.
\textbf{In-scope.} The agent roster implied by
\ref{fig:generator-stages}; vLLM TurboQuant routing for all LLM
calls; complexification step and the two critic agents.
\textbf{Out-of-scope.} Non-agentic synthesis paths.
\textbf{Paths to extend.} Generation engine in
\texttt{src/training/pipeline/}; vLLM client wrapper.
\textbf{Definition of done.} Running Algorithm~2 against a taxonomy
produces \texttt{TrainingExample} records that pass both critics and
validate against the training-example record schema.
\textbf{Verification.} Generation unit tests pass on fixture inputs;
double-critic agreement rate meets the threshold in
chapter~\ref{chap:evaluation}.

\subsubsection{Task S4 --- Environment and CLI.}
\textbf{Intent.} Deliver the environment and CLI surfaces specified
in chapter~\ref{chap:config}, including the configuration table, CLI
usage, and output files listed in~\ref{tab:outputs}.
\textbf{In-scope.} Configuration loader; CLI subcommands for each
pipeline stage; output-file writers.
\textbf{Out-of-scope.} IDE integrations.
\textbf{Paths to extend.} CLI module in
\texttt{src/training/pipeline/}.
\textbf{Definition of done.} Each configuration key in the
configuration table is read and respected; the CLI runs end-to-end
and produces every output file in~\ref{tab:outputs}.
\textbf{Verification.} CLI smoke tests pass; output files match the
schemas declared for them.

\subsubsection{Task S5 --- Evaluation framework.}
\textbf{Intent.} Deliver the evaluation framework in
chapter~\ref{chap:evaluation}, including calibrated complexity
scoring (Algorithm~3, Elo), taxonomic coverage evaluation,
embedding-based diversity, and critic-calibration validation.
\textbf{In-scope.} Each evaluator as a runnable function; report
writers.
\textbf{Out-of-scope.} Real-time evaluation dashboards beyond what
the chapter defines.
\textbf{Paths to extend.} Evaluation module in
\texttt{src/training/pipeline/}.
\textbf{Definition of done.} Each evaluator runs against a fixture
corpus and emits a report that validates against the corresponding
report schema.
\textbf{Verification.} Evaluator unit tests pass; reports reproduce
under fixed seeds.

\subsubsection{Task S6 --- Test fixtures and testing.}
\textbf{Intent.} Deliver the test fixtures specification in
chapter~\ref{ch:test-fixtures} and the testing specification in
chapter~\ref{ch:testing} at the coverage thresholds
in~\ref{tab:simula-coverage}.
\textbf{In-scope.} Invoice, OCR, taxonomy, and training-example
fixtures; unit tests for taxonomy traversal and generation; CI
wiring.
\textbf{Out-of-scope.} Performance benchmarking infrastructure.
\textbf{Paths to extend.} Test directory colocated with the Simula
pipeline code.
\textbf{Definition of done.} Coverage meets~\ref{tab:simula-coverage};
CI gates block regressions.
\textbf{Verification.} \texttt{make spec-simula} remains green; the
Simula test entry points all run.


\section{Governance obligations applied to this domain}
\label{sec:agent-governance}
The regulations domain wraps Simula at runtime. Four obligations
apply without exception; each is stated as a concrete hook in the
Simula codebase.

\begin{itemize}
  \item \textbf{MAS MGF four pillars} (see regulations
    chapter~3 in
    \texttt{docs/latex/specs/regulations/chapters/03-mgf-framework.tex}).
    Pillar~1 (assess and bound): every Simula MCP tool is declared
    in the allow-list in \S\ref{sec:agent-persona}, deny-by-default.
    Pillar~2 (design and deploy): the pipeline stages in
    chapter~\ref{chap:generator} are traceable end-to-end via the
    per-request identity attribution described below. Pillar~3
    (operate and monitor) and pillar~4 (govern and disclose) are
    satisfied by emitting the evidence records produced by the
    Simula evaluation framework (chapter~\ref{chap:evaluation}) into
    the shared governance store.
  \item \textbf{Identity attribution} (regulations chapter~10,
    \texttt{10-identity-attribution.tex}). Every MCP call and every
    pipeline run carries the persona id, agent id, request id, and
    tenant id. The gateway is the attachment point; Simula code
    propagates the identity envelope without mutating it.
  \item \textbf{Emergent-capability monitoring} (regulations
    chapter~8, \texttt{08-monitoring-spec.tex}). Extraction,
    taxonomy generation, synthesis, and evaluation steps are
    instrumented with the monitoring metrics specified in that
    chapter; metrics flow to the monitoring sink defined there.
  \item \textbf{Conformance tooling} (regulations chapter~6,
    \texttt{06-conformance-tooling.tex}). The Simula service
    exposes AI Verify 2.0 plug-in hooks and Moonshot CI/CD 1.1.0
    benchmark entry points as required by that chapter.
\end{itemize}

\section{Cross-spec integration --- full detail}
\label{sec:agent-integration}

\subsection{With Arabic AP}
Simula consumes Arabic's schemas and fixtures as training input. The
contract is schema-in, \texttt{TrainingExample}-out: Simula reads
\texttt{invoice.schema.json}, \texttt{vat-checklist.schema.json}, and
the Arabic fixture store; it produces \texttt{TrainingExample}
records. Simula does not call Arabic MCP tools at runtime and does
not share state with the Arabic workflow. No cross-domain
operational envelope is introduced.

\subsection{With Trial Balance}
Simula consumes TB's schemas and fixtures as training input. The
contract is schema-in, \texttt{TrainingExample}-out: Simula reads
the TB schema set and the TB fixture store; it produces
\texttt{TrainingExample} records. Simula does not call TB MCP tools
at runtime and does not share state with the TB workflow.

\subsection{With Regulations}
Regulations wraps Simula at runtime. Every Simula MCP call is
brokered through the regulations governance gate, which applies
pillars~1--4 of MAS MGF, attaches the identity envelope, and routes
emergent-capability metrics to the monitoring store. The Simula
service code does not call regulations directly; the gateway layer
enforces this. Simula also consumes regulations' schemas
(requirement, conformance-tool, monitoring metrics, identity and
audit) as training input.

\subsection{With Simula}
Simula exports the training-example record, the schema-registry
record, the taxonomy structure, and the enum vocabularies under
\texttt{structured/schema/} (chapter~\ref{chap:schema}); registers
the MCP tools enumerated in \S\ref{sec:agent-persona}; and applies
the four governance hooks in \S\ref{sec:agent-governance} to every
call it serves. External consumers see Simula only through these
contracts and through the corpus it emits; there are no private
back-channels.

\section{Wave assignment and sequencing}
\label{sec:agent-waves}
Implementation runs in three waves.

\begin{itemize}
  \item \textbf{Wave 1 --- foundations.} Unified schema registry
    entry; persona-gated surface for the Training Engineer;
    identity attribution envelope flowing through the gateway; HANA
    extraction pipeline (Task S1). Prerequisites: the regulations
    chapter~10 identity schema must be frozen.
  \item \textbf{Wave 2 --- Simula core build-out.} Tasks S2, S3, S4,
    S5. Prerequisites: Wave~1 complete; vLLM TurboQuant endpoint
    reachable; Arabic, TB, and regulations schema sets published.
  \item \textbf{Wave 3 --- hardening and conformance.} Task S6 and
    full conformance-tooling hook-up. Prerequisites: Wave~2
    complete; regulations Wave~2 conformance tooling integrated;
    monitoring sink live.
\end{itemize}

Simula occupies Wave~1 for its extraction pipeline (because other
domains depend on Simula's fixtures being reachable), Wave~2 for its
core build-out, and Wave~3 for its testing and conformance surfaces.

\section{Acceptance criteria (platform-wide)}
\label{sec:agent-acceptance}
When all four chapter~12 instruction sets are implemented, the
platform satisfies the following invariants:

\begin{enumerate}
  \item \textbf{One instance.} A single codebase runs a single
    logical instance; no domain ships as a separate product.
  \item \textbf{Persona-gated surfaces.} The four personas each
    reach a tailored UI on the shared shell; no route, MCP tool, or
    data product is visible outside its persona's entitlement set.
  \item \textbf{Unified registry.} Every schema in use is listed in
    the ODPS~4.1 registry under its owning domain; validators
    resolve \texttt{\$ref}s through the unified registry only.
  \item \textbf{No schema merges.} The two
    \texttt{requirement.schema.json} files in TB and regulations
    remain distinct; no cross-domain composite schemas are produced.
  \item \textbf{No cross-domain operational envelope.} Runtime
    orchestration stays within each domain's service; only Simula
    consumes other domains' outputs, and only as training input.
  \item \textbf{Governance wraps all three non-regulations domains.}
    Arabic, TB, and Simula service calls pass through the
    regulations runtime wrapper without exception.
\end{enumerate}

\section{References}
\label{sec:agent-references}
Within-spec references:
chapter~\ref{chap:overview} (overview),
chapter~\ref{chap:schema} (data schema),
chapter~\ref{chap:extraction} (HANA extraction pipeline),
chapter~\ref{chap:taxonomy} (taxonomy engine, Algorithm~1),
chapter~\ref{chap:generator} (generation engine, Algorithm~2),
chapter~\ref{chap:config} (environment and CLI),
chapter~\ref{chap:evaluation} (evaluation framework),
chapter~\ref{ch:test-fixtures} (test fixtures),
chapter~\ref{ch:testing} (testing specification),
chapter~\ref{sec:sim-roadmap} (project roadmap).

Peer chapter~12 instruction sets for the other three domains (each
is self-contained; consult directly when operating in that domain):
\begin{itemize}
  \item Arabic AP:
    \texttt{docs/latex/specs/arabic/chapters/12-implementation-instructions.tex}.
  \item Trial Balance:
    \texttt{docs/latex/specs/tb/chapters/12-implementation-instructions.tex}.
  \item Regulations:
    \texttt{docs/latex/specs/regulations/chapters/12-implementation-instructions.tex}.
\end{itemize}

Shared platform artefacts:
\begin{itemize}
  \item ODPS~4.1 registry:
    \texttt{src/intelligence/ai-core-pal/data\_products/registry.yaml}.
  \item Schema pipeline and resolver:
    \texttt{src/intelligence/schema\_pipeline/}.
  \item Training pipeline (Simula):
    \texttt{src/training/pipeline/}.
  \item Shared workspace shell:
    \texttt{generativeUI/training-webcomponents-ngx}.
\end{itemize}
