\chapter{Schema Versioning and Interface Contracts}
\label{chap:schema-versioning}

This chapter defines the schema versioning policy for the Simula training data
framework, ensuring stable cross-domain interfaces and controlled evolution of
data contracts.

\section{Versioning Policy}

\subsection{Semantic Versioning for Schemas}

All Simula schemas follow Semantic Versioning (SemVer) 2.0.0:

\begin{itemize}
  \item \textbf{MAJOR.MINOR.PATCH} (e.g., \texttt{1.2.0})
  \item \textbf{MAJOR}: Breaking changes to required fields or core structure
  \item \textbf{MINOR}: Backward-compatible additions (new optional fields)
  \item \textbf{PATCH}: Documentation, description, or metadata updates
\end{itemize}

\subsection{Current Schema Versions}

\begin{table}[htbp]
\centering
\caption{Simula schema versions as of v1.2.}
\label{tab:schema-versions}
\footnotesize
\begin{tabularx}{\textwidth}{@{}l l l X@{}}
\toprule
\textbf{Schema} & \textbf{Version} & \textbf{Status} & \textbf{Description} \\
\midrule
\texttt{training-example.schema.json} & 1.2.0 & Stable & Training example record with critic evaluation \\
\texttt{taxonomy.schema.json} & 1.2.0 & Stable & Taxonomy tree structure with quality metrics \\
\texttt{schema-registry.schema.json} & 1.1.0 & Stable & HANA schema registry record \\
\texttt{evaluation-report.schema.json} & 1.0.0 & Draft & Evaluation framework output \\
\bottomrule
\end{tabularx}
\end{table}

\section{Schema Evolution Rules}

\subsection{Backward Compatibility Requirements}

\begin{enumerate}
  \item \textbf{Required fields cannot be removed} in MINOR or PATCH versions.
  \item \textbf{Required fields cannot be renamed} without a MAJOR version bump.
  \item \textbf{New optional fields} can be added in MINOR versions.
  \item \textbf{Enum values can only be added}, not removed, in MINOR versions.
  \item \textbf{Removing enum values} requires a MAJOR version bump with deprecation notice.
\end{enumerate}

\subsection{Deprecation Process}

\begin{enumerate}
  \item Mark field as deprecated in schema description (MINOR version).
  \item Add \texttt{deprecated: true} metadata if supported by tooling.
  \item Document migration path in changelog.
  \item Maintain deprecated field for at least one MINOR release cycle.
  \item Remove deprecated field in next MAJOR version.
\end{enumerate}

\section{Cross-Domain Interface Contracts}

\subsection{Schema-In, TrainingExample-Out Contract}

Simula maintains a strict contract with consumer domains:

\begin{lstlisting}[basicstyle=\ttfamily\small,caption={Cross-domain interface contract.}]
Input (from Arabic AP, TB, Regulations):
  - Schema files conforming to each domain's schema registry
  - Fixture files for validation

Output (to consumers):
  - TrainingExample records conforming to training-example.schema.json v1.2.0
  - Taxonomy artifacts conforming to taxonomy.schema.json v1.2.0
  - Evaluation reports conforming to evaluation-report.schema.json v1.0.0
\end{lstlisting}

\subsection{Interface Versioning}

Each cross-domain interface is versioned independently:

\begin{table}[htbp]
\centering
\caption{Cross-domain interface versions.}
\label{tab:interface-versions}
\footnotesize
\begin{tabularx}{\textwidth}{@{}l l l X@{}}
\toprule
\textbf{Interface} & \textbf{Version} & \textbf{Consumer} & \textbf{Contract} \\
\midrule
Simula $\to$ TB & 1.2 & Trial Balance & TrainingExample output \\
Simula $\to$ Arabic & 1.2 & Arabic AP & TrainingExample output \\
Simula $\to$ Regulations & 1.2 & Regulations & TrainingExample output \\
TB $\to$ Simula & 1.0 & Simula & Schema + fixture input \\
Arabic $\to$ Simula & 1.0 & Simula & Schema + fixture input \\
Regulations $\to$ Simula & 1.0 & Simula & Schema + fixture input \\
\bottomrule
\end{tabularx}
\end{table}

\section{Schema Validation}

\subsection{Validation Points}

Schema validation is enforced at three points:

\begin{enumerate}
  \item \textbf{Input validation}: Before processing consumer schemas
  \item \textbf{Output validation}: Before emitting TrainingExample records
  \item \textbf{Report validation}: Before publishing evaluation reports
\end{enumerate}

\subsection{Validation Implementation}

\begin{lstlisting}[language=jsonx,caption={Schema validation in the pipeline.}]
from jsonschema import validate, ValidationError
import json

def validate_training_example(example: dict) -> bool:
    """Validate a training example against the schema."""
    with open("docs/schema/simula/training-example.schema.json") as f:
        schema = json.load(f)
    
    try:
        validate(instance=example, schema=schema)
        return True
    except ValidationError as e:
        logger.error(f"Schema validation failed: {e.message}")
        return False

# Usage in the pipeline
for example in generated_examples:
    if not validate_training_example(example.to_dict()):
        raise SchemaValidationError(f"Example {example.id} failed validation")
\end{lstlisting}

\section{Changelog Management}

\subsection{Schema Changelog Location}

Schema changes are documented in:
\begin{itemize}
  \item \texttt{docs/schema/CHANGELOG.md} --- consolidated schema changelog
  \item Each schema file's \texttt{description} field --- version-specific notes
\end{itemize}

\subsection{Changelog Entry Format}

\begin{lstlisting}[basicstyle=\ttfamily\small,caption={Changelog entry format.}]
## [1.2.0] - 2026-04-21

### Added
- `critic_evaluation` complex field for double-critic results
- `generation_metadata` for reproducibility tracking
- `quality_signals` for downstream analysis

### Changed
- `complexity_score` now required (was optional in 1.1.0)

### Deprecated
- `difficulty` field (use `complexity_level` instead)
- `taxonomy_mix` field (use `taxonomy_path` instead)

### Migration
- Existing producers should add `complexity_score` with default 0.5
- Replace `difficulty` with `complexity_level` (EASY, MEDIUM, HARD)
\end{lstlisting}

\section{Version Negotiation}

\subsection{Runtime Version Checking}

The pipeline checks schema versions at runtime:

\begin{lstlisting}[language=jsonx,caption={Version negotiation.}]
def check_schema_compatibility(
    schema_path: str,
    min_version: str = "1.0.0",
    max_version: str = "2.0.0",
) -> bool:
    """Check if schema version is within supported range."""
    with open(schema_path) as f:
        schema = json.load(f)
    
    schema_id = schema.get("$id", "")
    # Extract version from $id or metadata
    version = extract_version(schema_id)
    
    return semver.match(version, f">={min_version}") and \
           semver.match(version, f"<{max_version}")
\end{lstlisting}

\subsection{Graceful Degradation}

When encountering an unknown schema version:

\begin{enumerate}
  \item Log a warning with schema ID and version
  \item Attempt to process with latest known compatible schema
  \item Flag output with \texttt{schema\_version\_mismatch: true}
  \item Include warning in generation stats
\end{enumerate}

\section{Implementation Checklist}

\begin{itemize}
  \item[\texttt{[x]}] Schema version in \texttt{\$id} field
  \item[\texttt{[x]}] \texttt{training-example.schema.json} at v1.2.0
  \item[\texttt{[x]}] \texttt{taxonomy.schema.json} at v1.2.0
  \item[\texttt{[ ]}] Automated schema validation in CI
  \item[\texttt{[ ]}] Version negotiation in runtime
  \item[\texttt{[ ]}] Cross-domain contract tests
  \item[\texttt{[ ]}] Schema changelog automation
\end{itemize}

\section{Governance}

\subsection{Schema Change Approval}

Schema changes require:

\begin{enumerate}
  \item RFC submitted to Controls Architecture Team
  \item 5-day review period for MINOR changes
  \item 10-day review period for MAJOR changes
  \item All consuming spec owners must approve breaking changes
  \item One-release deprecation window for breaking changes
\end{enumerate}

\subsection{Change Ownership}

\begin{table}[htbp]
\centering
\caption{Schema ownership.}
\label{tab:schema-ownership}
\footnotesize
\begin{tabularx}{\textwidth}{@{}l l X@{}}
\toprule
\textbf{Schema} & \textbf{Owner} & \textbf{Approval Authority} \\
\midrule
\texttt{training-example.schema.json} & Simula Lead & Simula + Consumer Leads \\
\texttt{taxonomy.schema.json} & Simula Lead & Simula Lead \\
\texttt{common/enums.yaml} & Controls Arch & All Domain Leads \\
\bottomrule
\end{tabularx}
\end{table}