% =============================================================================
% Chapter 2 --- System Use Cases in Scope
% =============================================================================
\chapter{System Use Cases in Scope}
\label{ch:usecases}

\section{Use-case inventory purpose}

Every AI-enabled capability in the repository must be registered as a use case
before implementation begins. This chapter defines the operational use-case
inventory that is subject to governance controls.

\section{Operational use-case inventory}

\begin{longtable}{@{}l p{4cm} l l p{4.5cm}@{}}
\caption{Use-Case Inventory for Committee Review} \\
\toprule
\textbf{ID} & \textbf{Use Case} & \textbf{Action Type} & \textbf{Risk Tier} & \textbf{Core Risk Drivers} \\
\midrule
\endfirsthead
\multicolumn{5}{c}{\tablename\ \thetable{} -- continued} \\
\toprule
\textbf{ID} & \textbf{Use Case} & \textbf{Action Type} & \textbf{Risk Tier} & \textbf{Core Risk Drivers} \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{Continued on next page} \\
\endfoot
\bottomrule
\endlastfoot
\usecaseid{UC-01} & Chat and prompt-response synthesis & Read/generate & Medium & Prompt injection, hallucination, confidentiality \\
\usecaseid{UC-02} & AI-generated UI actions & Assist/approve & High & Automation bias, hidden writes \\
\usecaseid{UC-03} & HANA query and data-product prompts & Read/query & High & Schema exposure, SQL guardrails \\
\usecaseid{UC-04} & Retrieval and vector operations & Read/write vectors & High & Sensitive ingestion, retrieval leakage \\
\usecaseid{UC-05} & Personal knowledge and memory & Read/write knowledge & High & Retention, user boundaries \\
\usecaseid{UC-06} & OCR document extraction & Read/extract & High & Document sensitivity, interpretation \\
\usecaseid{UC-07} & Data-cleaning workflows & Recommend/process & Medium & False assurance, traceability \\
\usecaseid{UC-08} & Training and deployment & Create/deploy & Critical & Provenance, validation sufficiency \\
\usecaseid{UC-09} & Pipeline orchestration & Automate workflow & High & Process chaining, silent failure \\
\usecaseid{UC-10} & Agent/tool orchestration (MCP) & Tool use/plan & Critical & Tool misuse, dynamic planning \\
\usecaseid{UC-11} & Governance and audit & Oversight/evidence & High & Approval integrity, completeness \\
\usecaseid{UC-12} & Local model serving & Inference runtime & High & Confidential routing, resilience \\
\end{longtable}

\section{Use-case descriptions}

\subsection{UC-01: Chat and prompt-response synthesis}

\begin{description}
  \item[Description] Interactive chat interfaces where users submit prompts and
        receive LLM-generated responses.
  \item[Action Type] Read/generate --- reads context, generates new content.
  \item[Risk Tier] Medium
  \item[Primary Risks] Prompt injection, hallucination, confidentiality leakage.
  \item[Key Controls] \controlid{SEC-01}, \controlid{EVAL-01}, \controlid{USR-01}
\end{description}

\subsection{UC-02: AI-generated UI actions}

\begin{description}
  \item[Description] AI assists users by suggesting or executing UI actions
        based on context.
  \item[Action Type] Assist/approve --- proposes actions for human approval.
  \item[Risk Tier] High
  \item[Primary Risks] Automation bias, hidden writes to backend systems.
  \item[Key Controls] \controlid{HUM-01}, \controlid{HUM-02}, \controlid{AGT-01}
\end{description}

\subsection{UC-03: HANA query and data-product prompts}

\begin{description}
  \item[Description] Natural language queries translated to SQL against SAP HANA.
  \item[Action Type] Read/query --- reads data, executes queries.
  \item[Risk Tier] High
  \item[Primary Risks] Schema exposure, SQL injection, guardrail bypass.
  \item[Key Controls] \controlid{DATA-01}, \controlid{SEC-01}, \controlid{AGT-02}
\end{description}

\subsection{UC-08: Training and deployment}

\begin{description}
  \item[Description] Model training, fine-tuning, and deployment to production.
  \item[Action Type] Create/deploy --- creates new model artefacts.
  \item[Risk Tier] Critical
  \item[Primary Risks] Training data provenance, validation sufficiency.
  \item[Key Controls] \controlid{GOV-03}, \controlid{EVAL-01}, \controlid{EVAL-03}
\end{description}

\subsection{UC-10: Agent/tool orchestration (MCP)}

\begin{description}
  \item[Description] Multi-step agent planning with tool invocation via MCP.
  \item[Action Type] Tool use/plan --- invokes external tools and services.
  \item[Risk Tier] Critical
  \item[Primary Risks] Tool misuse, dynamic planning errors, unbounded autonomy.
  \item[Key Controls] \controlid{AGT-01}, \controlid{AGT-02}, \controlid{AGT-03},
        \controlid{AGT-04}, \controlid{HUM-01}, \controlid{MON-01}
\end{description}

\section{Risk tier definitions}

\begin{table}[htbp]
\centering
\caption{Risk Tier Definitions}
\begin{tabularx}{\textwidth}{@{}l l X@{}}
\toprule
\textbf{Tier} & \textbf{IRS Range} & \textbf{Description} \\
\midrule
Low & $0 \leq IRS < 25$ & Minimal risk, standard controls \\
Medium & $25 \leq IRS < 50$ & Moderate risk, enhanced controls \\
High & $50 \leq IRS < 75$ & Significant risk, critical controls required \\
Critical & $75 \leq IRS \leq 100$ & Highest risk, full control suite and review board \\
\bottomrule
\end{tabularx}
\end{table}

\section{Use-case registration process}

Before implementation, each use case must be registered with:

\begin{enumerate}
  \item \textbf{Use-case card} --- structured description including ID, name,
        action type, data classification, autonomy level
  \item \textbf{Risk assessment} --- calculated Inherent Risk Score (IRS)
  \item \textbf{Control profile} --- applicable controls identified
  \item \textbf{Owner assignment} --- named business and technical owners
  \item \textbf{Ticket reference} --- tracking ticket for governance approval
\end{enumerate}

Registration is tracked in the use-case section of the regulations
requirement corpus at
\filepath{docs/regulations/machine-readable/specification-addendum.yaml}
(key \texttt{use\_cases}), validated against the schemas under
\filepath{docs/schema/regulations/} via
\filepath{docs/schema/validate.py}.