% !TEX program = xelatex
% -----------------------------------------------------------------------------
% Arabic AP: Business Requirements and Lineage Specification
% -----------------------------------------------------------------------------
\documentclass[11pt,a4paper,twoside]{report}

\usepackage{xcolor}
\usepackage[a4paper,margin=28mm]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{tcolorbox}
\usepackage[table]{colortbl}

% --- SAP Midnight Palette ----------------------------------------------------
\definecolor{sapblue}{HTML}{0B5FA5}
\definecolor{sapmidnight}{HTML}{1C2B39}
\definecolor{sapgrey}{HTML}{546E7A}
\definecolor{sapamber}{HTML}{B8360F}

% --- Semantic Hyperlinking --------------------------------------------------
\usepackage[hidelinks]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}

% --- Safe Font Fallback (Bypassing polyglossia font errors) ------------------
\usepackage{fontspec}
\setmainfont{Arial}
\newcommand{\ar}[1]{#1}

\newtcolorbox{adrbox}[2][]{colback=blue!5,colframe=blue!75!black,fonttitle=\bfseries,title=ADR: #2,#1}

\begin{document}

\title{\Huge Arabic AP Domain \\
       \textbf{Business Requirements \& Lineage} \\
       \large Source-to-Specification Mapping \\
       \small Ref: \texttt{ARA-REQ-LIN-2026-v1.0}}
\author{SAP-OSS Business Architecture Team}
\date{April 2026}

\maketitle

\chapter{Source Material Registry}
\label{ch:source-registry}

This chapter catalogues the primary source documents used to derive the business requirements for the Arabic AP domain.

\begin{longtable}{p{5cm} p{3cm} p{6cm}}
\caption{Source Document Inventory} \\
\toprule
\textbf{File Name} & \textbf{Regulatory Body} & \textbf{Primary Concern} \\
\midrule
\endfirsthead
\toprule
\textbf{File Name} & \textbf{Regulatory Body} & \textbf{Primary Concern} \\
\midrule
\endhead
\texttt{Accounts Payable DOI.docx} & SCB Bahrain / NBR & Delegation of Invoicing (DOI) \& Controls \\
\texttt{Saudi - OTP Entry processing...} & ZATCA (Saudi) & VAT Q3'25 processing rules \\
\texttt{EDAA Annual Fee Invoice...} & EDAA & Arabic/English Invoice layout (Standard) \\
\texttt{Arabic-Simplified-Invoice.pdf} & Local Authority & Arabic-only simplified invoice structure \\
\bottomrule
\end{longtable}

\chapter{Transformation Logic}
\label{ch:transformation}

\section{From PDF to Machine-Readable (YAML/JSON)}
The business requirements were extracted from the sources in \cref{ch:source-registry} using a semi-automated "Ground Truth" process:
\begin{enumerate}
    \item \textbf{Heuristic Extraction:} Initial requirement identification by GFS Analysts.
    \item \textbf{Schema Normalization:} Mapping requirements to the \texttt{invoice.schema.json} and \texttt{vat-checklist.schema.json}.
    \item \textbf{Formalization:} Converting narrative rules (e.g., "Full tax invoice must contain TIN") into JSON-Logic predicates.
\end{enumerate}

\chapter{Business Requirements Catalogue}
\label{ch:requirements}

\begin{longtable}{p{2cm} p{5cm} p{7cm}}
\caption{Consolidated Business Requirements} \\
\toprule
\textbf{Req ID} & \textbf{Business Rule} & \textbf{Source Document Reference} \\
\midrule
\endfirsthead
\toprule
\textbf{Req ID} & \textbf{Business Rule} & \textbf{Source Document Reference} \\
\midrule
\endhead
\rowcolor{sapblue!5}
\textbf{BR-OCR-01} & Character precision for currency must exceed 99.5\%. & \texttt{Saudi - OTP Entry processing...} (p. 2) \\
\textbf{BR-VAT-SA} & Simplified vs Full Invoice determination based on 1000 SAR threshold. & ZATCA VAT Guidelines (Implicit in Saudi OTP) \\
\rowcolor{sapblue!5}
\textbf{BR-DOI-07} & Evidence-based seven-control sign-off for Bahrain. & \texttt{Accounts Payable DOI.docx} (Section 4) \\
\textbf{BR-TRN-01} & Translation must maintain semantic mapping for legal entity names. & \texttt{EDAA Annual Fee Invoice...} \\
\bottomrule
\end{longtable}

\chapter{Requirement-to-Specification Lineage}
\label{ch:lineage}

\begin{longtable}{p{2cm} p{6cm} p{6cm}}
\caption{Technical Traceability Matrix} \\
\toprule
\textbf{Req ID} & \textbf{Technical Implementation} & \textbf{Implementation Spec Ref} \\
\midrule
\rowcolor{sapamber!5}
\textbf{BR-OCR-01} & Confidence threshold calibration pilot. & \texttt{arabic-ap-spec} Ch. 10 \\
\textbf{BR-VAT-SA} & Rule ID \texttt{VAT-SA-SI-001} in VAT Engine. & \texttt{arabic-ap-spec} Ch. 8 \\
\rowcolor{sapamber!5}
\textbf{BR-DOI-07} & 28-state workflow wiring in \texttt{ap.workflow.advance}. & \texttt{arabic-ap-spec} Ch. 5 \& 6 \\
\textbf{BR-TRN-01} & Arabic--English translation gate calibration. & \texttt{arabic-ap-spec} Ch. 3 \\
\bottomrule
\end{longtable}

\end{document}
