\chapter{Canonical data schema}\label{ch:schema}

\section{Layout of \texttt{structured/}}

The machine-readable corpus is organised as in \cref{tab:layout}.
All records validate against a JSON Schema Draft~07 file under
\corpus\texttt{structured/schema/}. The
\corpus\texttt{structured/validate.py} script walks the tree, resolves
\texttt{\$ref} entries via a local registry (so offline validation
succeeds), and exits non-zero on any failure; it is the authoritative
acceptance gate for this chapter.

\begin{table}[htbp]
\centering
\caption{Directory layout under \corpus\texttt{structured/}.}
\label{tab:layout}
\small
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Path} & \textbf{Contents} \\ \midrule
\texttt{schema/invoice.schema.json}            & Canonical invoice / assessment record. \\
\texttt{schema/vat-checklist.schema.json}      & Regulator checklist (KSA 11-point; extensible). \\
\texttt{schema/payment-instruction.schema.json}& Vendor payment memo. \\
\texttt{schema/approval-workflow.schema.json}  & Declarative state machine. \\
\texttt{schema/doi-policy.schema.json}         & DOI with activities and controls. \\
\texttt{invoices/*.json}                       & Invoice records (EG ETA ×2; SA ZATCA ×1). \\
\texttt{payment-instructions/*.json}           & Qatar QCSD/EDAA records. \\
\texttt{vat-checklists/*.json}                 & Saudi OTP/VAT Q3 2025. \\
\texttt{policies/*.json}                       & Bahrain AP DOI. \\
\texttt{workflows/*.json}                      & PO-eProc, manual-one-off, manual-recurring, VAT payment. \\
\texttt{enums.yaml}                            & Controlled vocabulary. \\
\texttt{corpus.json}                           & Flat index of every source $\leftrightarrow$ text $\leftrightarrow$ record link. \\
\texttt{validate.py}                           & Offline validator. \\ \bottomrule
\end{tabularx}
\end{table}

\section{Invoice record}

The invoice record is the central object shared by ingestion (\cref{ch:ai}),
the VAT engine (\cref{ch:vat}), and the approval engine
(\cref{ch:ap}). \Cref{tab:invoice-fields} enumerates the top-level
fields; \cref{lst:invoice-example} is a worked example taken verbatim
from \corpus\path{structured/invoices/eg-eta-3CPJR5CQ4PNNN9A4XHBB370K10.json}.

\begin{table}[htbp]
\centering
\caption{Top-level fields of \texttt{invoice.schema.json}.}
\label{tab:invoice-fields}
\small
\begin{tabularx}{\linewidth}{@{}l l c X@{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Req.} & \textbf{Notes} \\ \midrule
\texttt{id}               & string & \checkmark & Filename stem under \texttt{invoices/}. \\
\texttt{metadata}         & object & \checkmark & Country, authority, dates, language, status. \\
\texttt{seller}           & party  & \checkmark & Includes Arabic name, English name, \texttt{tax\_id\_type}. \\
\texttt{buyer}            & party  & \checkmark & Same shape as seller. \\
\texttt{line\_items[]}    & array  & --         & At least one of \texttt{description}, \texttt{description\_ar}, \texttt{description\_en}. \\
\texttt{totals}           & object & \checkmark & \texttt{currency}, \texttt{gross} required. \\
\texttt{tax\_period}      & object & --         & Populated for periodic assessments (ZATCA). \\
\texttt{signature}        & object & --         & Kind $\in$ \{\textsc{seal}, \textsc{wet\_ink}, \textsc{digital}, \textsc{system\_generated}\}. \\
\texttt{source\_document} & object & \checkmark & \texttt{path}, \texttt{sha256}, \texttt{ocr\_text\_path}. \\
\texttt{notes[]}          & array  & --         & Free-form analyst notes. \\
\bottomrule
\end{tabularx}
\end{table}

\begin{figure}[htbp]
\begin{lstlisting}[language=jsonx,caption={Full invoice record for the
Egypt ETA seal invoice (\texttt{eg-eta-3CPJR5CQ4PNNN9A4XHBB370K10}).
Arabic strings are UTF-8; the LaTeX engine renders them RTL.},
label=lst:invoice-example]
{
  "id": "eg-eta-3CPJR5CQ4PNNN9A4XHBB370K10",
  "schema_ref": "invoice.schema.json",
  "metadata": {
    "country": "EG",
    "tax_authority": "ETA",
    "document_kind": "E_INVOICE",
    "electronic_id": "3CPJR5CQ4PNNN9A4XHBB370K10",
    "issue_date": "2025-07-15",
    "submission_date": "2025-07-15T16:10:00+00:00",
    "language": "ar",
    "status": "VALID"
  },
  "seller": {
    "name_ar": "محمد شريف شرف الدين",
    "tax_id": "438212282",
    "tax_id_type": "ETA_TRN",
    "activity_code": "1811"
  },
  "buyer": {
    "name_ar": "بنك ستاندرد تشارترد",
    "tax_id": "670307459",
    "tax_id_type": "ETA_TRN"
  },
  "line_items": [{
    "description_ar": "ختم ترودات دائرى 42 مللى سريلة",
    "description_en": "Circular Trodat seal 42mm, series",
    "quantity": 1, "unit_of_measure": "EA",
    "unit_price": 580000.00, "line_total": 580000.00
  }],
  "totals": {
    "currency": "EGP",
    "net": 580000.00, "vat_amount": 81200.00,
    "vat_rate_percent": 14, "vat_category": "FI",
    "gross": 661200.00
  },
  "source_document": {
    "path": "docs/arabic/source/3CPJR5CQ4PNNN9A4XHBB370K10.pdf",
    "sha256": "2b9126...4b2c0",
    "ocr_text_path": "docs/arabic/extracted-text/eg-eta-3CPJR...txt"
  }
}
\end{lstlisting}
\end{figure}

Invoice \textsc{electronic\_id} for ETA must match the regex in
\corpus\texttt{structured/enums.yaml}
(\verb|^[0-9A-Z]{26}$|); the ZATCA assessment number is a decimal prefix
matching the taxpayer's \textsc{zatca\_tdn}. The \texttt{status} enum is
\{\textsc{valid}, \textsc{invalid}, \textsc{pending}, \textsc{cancelled}\};
ZATCA assessments are always \textsc{valid} on arrival.

\section{VAT compliance checklist}

The KSA OTP/VAT checklist is encoded as an ordered array of items
(\cref{tab:checklist-fields}). Each item records the regulator's
question, the analyst's verdict (\textsc{y}/\textsc{n}/\textsc{na}), and
the extracted value. The authoritative list is the eleven points in the
Saudi Q3 2025 checklist; NBR (Bahrain) maintains an analogous set drawn
from Activity~3 of the DOI.

\begin{table}[htbp]
\centering
\caption{Fields of a VAT checklist record.}
\label{tab:checklist-fields}
\small
\begin{tabularx}{\linewidth}{@{}l l c X@{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Req.} & \textbf{Notes} \\ \midrule
\texttt{period}                 & object & \checkmark & \texttt{start}, \texttt{end}, \texttt{label}. \\
\texttt{checklist[]}            & array  & \checkmark & Items indexed 1\dots$n$; each has \texttt{result} $\in$ \{Y,N,NA\}. \\
\texttt{gl\_coding}             & object & \checkmark & Chart fields (account/product/dept/OU/class). \\
\texttt{supplier}               & object & \checkmark & \texttt{supplier\_type} $\in$ \{\textsc{government}, \textsc{vendor}, \textsc{employee}, \textsc{inter\_bu}\}. \\
\texttt{tax\_type\_classification} & string & \checkmark & VAT category enum. \\
\texttt{withholding\_tax}       & object & --         & Rate and amount if applicable. \\
\texttt{psid}                   & string & --         & Preparer PSID. \\
\bottomrule
\end{tabularx}
\end{table}

\begin{figure}[htbp]
\begin{lstlisting}[language=jsonx,caption={Excerpt from
\texttt{structured/vat-checklists/sa-otp-vat-q3-2025.json}
showing items 1--5 of the KSA 11-point checklist.},label=lst:checklist-excerpt]
"checklist": [
  {"n": 1, "label": "Invoice in Arabic language",                   "result": "Y"},
  {"n": 2, "label": "Addressed to Standard Chartered Capital",      "result": "Y"},
  {"n": 3, "label": "Date of issue",                                "result": "Y", "value": "2025-10-30"},
  {"n": 4, "label": "Sequential number",                            "result": "Y", "value": "3003093648253002"},
  {"n": 5, "label": "Tax Identification Number (15-digit)",         "result": "NA",
     "regulation": "Supplier is ZATCA itself -- government body"}
],
\end{lstlisting}
\end{figure}

\section{Payment instruction}

Derived from the Qatar QCSD instruction. The approval chain is an array
of typed signatories; each entry carries PSID, role, and approval
timestamp --- giving a tamper-evident audit trail once hashed.
The \texttt{chart\_field} is stored as a dash-separated composite
string (e.g.\ \texttt{665011-4551-800} = account-product-operating\_unit)
because this is the convention used by the source memos.

\section{Approval workflow}

Each workflow record is a deterministic state machine. States are nodes;
transitions are edges tagged with a trigger and, where relevant, a
\texttt{required\_role} and a list of \texttt{preconditions}. The
\texttt{delegation\_of\_authority} array gates roles by LCY threshold and
is consulted by the approval engine before any transition out of a
``$x$-approved'' state. Four workflow instances are shipped in the corpus:

\begin{itemize}
  \item \texttt{po-eproc} --- PO / eProcurement path (DOI Activity~3);
  \item \texttt{manual-one-off} --- one-off manual payment (DOI Activity~5);
  \item \texttt{manual-recurring} --- recurring manual with dispensation (DOI Activity~5);
  \item \texttt{vat-payment} --- authority-counterparty (e.g.\ ZATCA) path.
\end{itemize}

\section{DOI policy}

The Bahrain Accounts Payable DOI is represented as a
\texttt{doi-policy} record whose \texttt{activities[]} array preserves
the original numbering (1--7) and whose \texttt{controls[]} array gives
each control an ID, description, pointer back to the activity, and the
risk it mitigates. This is what the approval engine consults to decide
\emph{which} workflow instance applies to a given invoice.

\section{Enumerations}

The file \corpus\texttt{structured/enums.yaml} is the single source of
truth for controlled vocabulary used across the corpus: country codes,
tax-authority codes and portal URLs, VAT category codes,
\texttt{tax\_id\_type} regexes, business-unit codes, approval-role
codes, payment methods, and the five AI-session phase codes. Any new
enum value must be added to \texttt{enums.yaml} \emph{before} being
used in a schema or record; \texttt{validate.py} enforces this.

\section{Provenance}

Each record stores the SHA-256 of its source document in
\texttt{source\_document.sha256}. Reproducing the corpus from the
originals is therefore a single command:
\begin{lstlisting}[language=bash,basicstyle=\ttfamily\small]
find docs/arabic/source -type f -print0 \
  | xargs -0 shasum -a 256 \
  | diff - <(jq -r '.items[] | select(.sha256) | "\(.sha256)  \(.source_path)"' \
              docs/arabic/structured/corpus.json | sed 's|docs/arabic/source/||')
\end{lstlisting}
A clean diff proves the corpus is unaltered since the last
specification snapshot.

\section{Unified Schema Registry}

As of 18 April 2026, all schemas in this specification are published in the
unified schema registry at \path{docs/schema/}. Migration status is
\emph{complete} (see \path{docs/schema/registry.json} field
\texttt{migration.status}). This consolidation addresses cross-document schema
divergence identified in the specification review.

\begin{description}
  \item[Registry index.] \path{docs/schema/registry.json} --- Master index of
        all schemas across domains.
  \item[Common types.] \path{docs/schema/common/base-types.schema.json} ---
        Shared definitions (\texttt{Metadata}, \texttt{Currency},
        \texttt{TaxIdentifier}, \texttt{EntityParty}) referenced via JSON
        Schema \texttt{\$ref}.
  \item[Consolidated enums.] \path{docs/schema/common/enums.yaml} --- Single
        source of truth for enumeration values including \texttt{countries},
        \texttt{vat\_category}, \texttt{supplier\_type}, and
        \texttt{vat\_thresholds} used by this specification.
  \item[Arabic schemas.] \path{docs/schema/arabic/} --- Domain-specific
        schemas: \texttt{invoice.schema.json}, \texttt{vat-checklist.schema.json},
        \texttt{vendor.schema.json}, and \texttt{ocr-result.schema.json}.
        All marked \texttt{status: stable}.
\end{description}

\noindent
The per-country VAT thresholds (e.g.\ BHD~500 for Bahrain SI/FI boundary)
are defined in \path{docs/schema/common/enums.yaml\#/vat\_thresholds}
rather than in domain-specific \texttt{enums.yaml} files. Validation:

\begin{lstlisting}[language=bash]
python3 docs/schema/validate.py --registry-check
python3 docs/schema/validate.py --domain arabic \
    --schema invoice.schema.json --file invoice.json
\end{lstlisting}

\noindent
CI runs \texttt{--registry-check} on every pull request touching
\path{docs/schema/} or any dependent pipeline module.
