\chapter{Controls, testing, and acceptance}\label{ch:controls}

\section{Control catalogue}

The Bahrain DOI's seven controls
(\corpus\path{structured/policies/bh-accounts-payable-doi.json},
\texttt{controls[]}) are reproduced in \cref{tab:controls} with the
engine mechanism that enforces each.

\begin{table}[htbp]
\centering
\caption{Control catalogue and enforcement mechanisms.}
\label{tab:controls}
\small
\begin{tabularx}{\linewidth}{@{}l l X X@{}}
\toprule
\textbf{ID} & \textbf{Activity} & \textbf{Description} & \textbf{Enforced by} \\ \midrule
C1 & 4 & All required approvals present before payment release. & Workflow engine cannot transition to \textsc{paid} until every required-role approval has fired. \\
C2 & 5 & Vendor screening revalidated within 48~hours of payment. & Precondition on \texttt{cfo\_approved}$\to$\textsc{screened} transition (\texttt{manual-one-off}). \\
C3 & 4 & Amount payable within DOA per GDAM. & Linear DOA scan before any approval transition. \\
C4 & 5 & Dispensation in place for recurring manual payments. & Precondition on \textsc{draft}$\to$\textsc{dispensation\_ok} in \texttt{manual-recurring}. \\
C5 & 7 & Monthly PSGL$\leftrightarrow$eBBS reconciliation cleared within aging thresholds. & Batch job; escalation grid e-mails DBU FM / CFO. \\
C6 & 3 & FCY invoices revalued at central-bank rate on invoice date. & VAT engine runs FCY$\to$LCY \emph{before} DOA check. \\
C7 & 4 & Duplicate-payment check. & Intake Stage~1 short-circuits on matching \texttt{source\_document.sha256}. \\
\bottomrule
\end{tabularx}
\end{table}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  node distance=5mm and 6mm,
  stage/.style={draw,rounded corners,minimum width=22mm,minimum height=8mm,align=center,font=\scriptsize,fill=orange!10},
  ctrl/.style={draw,ellipse,minimum width=10mm,minimum height=6mm,font=\tiny,fill=red!10},
  t/.style={-{Latex[length=1.8mm]}},
  c/.style={-{Latex[length=1.6mm]},dashed,red!60!black}
]
\node[stage] (intake) {Intake};
\node[stage,right=of intake]  (ocr)  {OCR};
\node[stage,right=of ocr]     (ext)  {Extract};
\node[stage,right=of ext]     (vat)  {VAT engine};
\node[stage,below=of vat]     (doa)  {DOA scan};
\node[stage,left=of doa]      (apr)  {Approval};
\node[stage,left=of apr]      (pay)  {Pay};
\node[stage,left=of pay]      (rec)  {Reconcile};

\draw[t] (intake) -- (ocr);
\draw[t] (ocr) -- (ext);
\draw[t] (ext) -- (vat);
\draw[t] (vat) -- (doa);
\draw[t] (doa) -- (apr);
\draw[t] (apr) -- (pay);
\draw[t] (pay) -- (rec);

\node[ctrl,above=2.5mm of intake] (c7)  {C7};
\node[ctrl,above=2.5mm of vat]    (c6)  {C6};
\node[ctrl,above=2.5mm of doa]    (c3)  {C3};
\node[ctrl,below=2.5mm of apr]    (c1)  {C1};
\node[ctrl,below=2.5mm of pay]    (c2)  {C2};
\node[ctrl,below=2.5mm of pay,xshift=-14mm] (c4) {C4};
\node[ctrl,below=2.5mm of rec]    (c5)  {C5};

\draw[c] (c7) -- (intake);
\draw[c] (c6) -- (vat);
\draw[c] (c3) -- (doa);
\draw[c] (c1) -- (apr);
\draw[c] (c2) -- (pay);
\draw[c] (c4) -- (pay);
\draw[c] (c5) -- (rec);
\end{tikzpicture}
\caption{Control enforcement points along the end-to-end invoice
         lifecycle. Dashed arrows show where each DOI control
         (\cref{tab:controls}) is evaluated.}
\label{fig:controls-map}
\end{figure}

\section{Testing strategy}

Testing follows the five-session engagement model
(\cref{tab:sessions}). For each country rollout:

\begin{description}
  \item[Session 1 --- Understand.] Collect $\geq$5 sample invoices
        (regulator-issued if available), normalise into the corpus,
        confirm per-country ruleset is complete.
  \item[Session 2 --- Sketch.] Walk the end-to-end pipeline on paper
        using one sample invoice. Identify OCR pitfalls
        (\cref{tab:pitfalls}) specific to the country.
  \item[Session 3 --- Decide.] Freeze the rule precedence and the
        decision function for the VAT engine
        (\S\ref{ch:vat}).
  \item[Session 4 --- Prototype.] Run the full pipeline on the cohort;
        review each extracted record with the local finance team.
        Human-approval rate target $\geq$~0.75.
  \item[Session 5 --- Testing.] Independent run with $\geq$~5
        stakeholders; target $\geq$~0.90 human-approval with
        zero-edit rate measured.
\end{description}

\section{Regression golden set}

The three Arabic records already in
\corpus\texttt{structured/invoices/} (Egypt ETA $\times$ 2, Saudi
ZATCA $\times$ 1) together with the KSA checklist and the two Qatar
payment instructions constitute the regression golden set.

\paragraph{Invariants.} Every change to the pipeline must preserve:
\begin{enumerate}
  \item byte-identical \texttt{source\_document.sha256} for every
        corpus record (source files are immutable);
  \item 100\,\% schema validity under \texttt{validate.py};
  \item identical \texttt{tax\_type\_classification} for every
        invoice record;
  \item identical selected workflow for every record.
\end{enumerate}

\paragraph{Runner.} \texttt{validate.py} plus the following
\texttt{jq} equality check is the regression harness:

\begin{lstlisting}[language=bash,basicstyle=\ttfamily\small]
# Per-field equality of every golden-set record between a reference
# snapshot and the pipeline output.
diff \
  <(jq -S . docs/arabic/structured/invoices/*.json) \
  <(for p in docs/arabic/source/*.pdf; do
      ./pipeline extract "$p"
    done | jq -S .)
\end{lstlisting}

A non-empty diff is a regression.

\section{Acceptance criteria per workstream}

\paragraph{AI pipeline (\cref{ch:ai}).}
\begin{itemize}
  \item Corpus integrity check passes (every \texttt{corpus.json} path
        exists; every SHA-256 matches).
  \item For every source PDF: OCR confidence $\geq$~30\% on every page,
        schema-valid record emitted.
  \item Translation back-translation distance $\leq$~0.30 on every
        non-empty \texttt{description\_en}.
\end{itemize}

\paragraph{VAT engine (\cref{ch:vat}).}
\begin{itemize}
  \item KSA: all 11 checklist rules evaluated; decision function
        emits expected VAT category for every golden invoice.
  \item Egypt: \texttt{EXEMPT\_ACTIVITIES} lookup table complete for
        activities present in the corpus.
  \item Bahrain: each of the 11 NBR invoice types has a corresponding
        enum code.
\end{itemize}

\paragraph{AP engine (\cref{ch:ap}).}
\begin{itemize}
  \item Every workflow record loadable and schema-valid.
  \item DOA scan deterministic under round-trip FX conversion.
  \item Workflow-selection decision tree (\cref{fig:workflow-select})
        yields the documented workflow for every corpus record.
\end{itemize}

\section{Traceability matrix}

\Cref{tab:trace} links each specification section to the source
document that justifies it and to the structured record that realises
it. Every row must be present for a full audit pass.

\begin{table}[htbp]
\centering
\caption{Traceability of specification to source document and
         structured record.}
\label{tab:trace}
\footnotesize
\begin{tabularx}{\linewidth}{@{}l X X X@{}}
\toprule
\textbf{Section} & \textbf{Requirement} & \textbf{Source} & \textbf{Structured record} \\ \midrule
\cref{ch:overview} & Countries in scope & Manoj e-mail, 2025-11-23 & \path{enums.yaml} (countries) \\
\cref{ch:schema}   & Canonical invoice fields & EG ETA PDFs, ZATCA PDF & \path{schema/invoice.schema.json} \\
\cref{ch:schema}   & 11-point checklist shape & Saudi OTP/VAT Q3 2025 & \path{schema/vat-checklist.schema.json} \\
\cref{ch:ai}       & OCR command line & --- (new) & \path{extracted-text/*.txt} \\
\cref{ch:ai}       & Translation glossary & \ar{اشعار صدور فاتورة} (005).pdf & \path{invoices/sa-zatca-ishaar-005.json} \\
\cref{ch:vat}      & KSA rules (1--11) & Saudi OTP/VAT Q3 2025 & \path{vat-checklists/sa-otp-vat-q3-2025.json} \\
\cref{ch:vat}      & ETA 26-char ID regex & EG ETA PDFs & \path{enums.yaml} (tax\_authorities.ETA) \\
\cref{ch:vat}      & NBR 11 invoice types & AP DOI Activity 3 & \path{policies/bh-accounts-payable-doi.json} \\
\cref{ch:ap}       & Activities 1--7 & AP DOI & \path{policies/bh-accounts-payable-doi.json} \\
\cref{ch:ap}       & \texttt{po-eproc} state machine & AP DOI Activity 3 & \path{workflows/po-eproc.json} \\
\cref{ch:ap}       & \texttt{manual-one-off} state machine & AP DOI Activity 5 (one-off) & \path{workflows/manual-one-off.json} \\
\cref{ch:ap}       & \texttt{manual-recurring} state machine & AP DOI Activity 5 (recurring) & \path{workflows/manual-recurring.json} \\
\cref{ch:ap}       & \texttt{vat-payment} state machine & KSA Q3 2025 flow + ZATCA assessment & \path{workflows/vat-payment.json} \\
\cref{ch:controls} & Controls C1--C7 & AP DOI + KSA checklist + observations & \path{policies/bh-accounts-payable-doi.json} \\
\bottomrule
\end{tabularx}
\end{table}

\section{Operational readiness}

A country is declared \emph{production-ready} once:
\begin{enumerate}
  \item the regression golden set runs clean for 30 consecutive days
        in the \textsc{uat} environment;
  \item Session~5 acceptance meets the human-approval-rate target
        (\cref{ch:ai});
  \item every VAT rule for the country is present in the ruleset and
        has at least one passing test case;
  \item the country CFO signs off on the workflow configuration,
        recorded as a \texttt{CFO} role entry on the \texttt{vat-payment}
        or \texttt{manual-one-off} workflow golden record.
\end{enumerate}

\section{Deployment and rollback}

\paragraph{Environments.} Three are in scope:
\begin{description}
  \item[\textsc{dev}.] Developer workstations and CI; reads the corpus
        from the Git working tree. No live invoices.
  \item[\textsc{uat}.] Shared test environment reading a frozen copy of
        the corpus; live DOI activity stops at \textsc{cfo\_approved}
        (no SAMA/SWIFT rail). Country finance and GBS APU run Session
        4/5 here.
  \item[\textsc{prod}.] Live environment; the only one authorised to
        transition to \textsc{paid}.
\end{description}

\paragraph{Sign-off matrix.} Each promotion requires a two-party
approval recorded as an event in the \texttt{events/} stream
(\cref{ch:ap}):

\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Transition} & \textbf{Approvers} & \textbf{Evidence} \\ \midrule
\textsc{dev}$\to$\textsc{uat}  & Tech lead + QA lead           & Green CI build, clean regression. \\
\textsc{uat}$\to$\textsc{prod} & Country CFO + GBS APU head    & Session 5 acceptance report; 30-day UAT stability; CFO workflow golden record. \\
\bottomrule
\end{tabularx}

\paragraph{Rollback.} Deployments are blue/green. The engine, rules,
and schemas are versioned with a content-addressed tag (SHA-256 of
\texttt{structured/corpus.json} at release time). Rollback is a
one-command revert to the previous tag; no database migration is
required because the canonical corpus is append-only. Any in-flight
invoice whose workflow record references a schema version that was
rolled back is automatically held in \textsc{awaiting\_review}.

\paragraph{Change control.} Ruleset edits (\cref{ch:vat}) and
workflow edits (\cref{ch:ap}) are routed through the same two-party
sign-off; schema edits require an additional sign-off from the AI
pipeline owner because they invalidate the regression golden set
(\S\ref{ch:ai}).

\section{Audit event log}

The approval engine emits an append-only event per transition
(\texttt{workflow\_id}, \texttt{from}, \texttt{to},
\texttt{actor\_psid}, \texttt{timestamp}, \texttt{evidence\_refs}),
hashing its predecessor so the stream is tamper-evident. The stream
is \emph{in scope} for compliance purposes but the storage substrate
(append-only object store, retention policy, restoration runbook) is
owned by the GBS APU audit team and specified in
\emph{SCB-APU-AUDIT-001}, referenced but not reproduced here. The
engine guarantees:

\begin{itemize}
  \item every workflow transition in \cref{ch:ap} emits exactly one
        event;
  \item events are linearly hash-chained;
  \item the event stream is the sole source of truth for DOA
        compliance (controls C1--C4 in \cref{tab:controls}).
\end{itemize}
