\chapter{Testing Specification}\label{ch:testing}

This chapter specifies the testing requirements for the Arabic AP invoice
processing pipeline, including test cases, coverage targets, and CI/CD
integration.

\section{Test coverage requirements}

\begin{table}[htbp]
\centering
\caption{Arabic AP test coverage targets.}
\label{tab:arabic-coverage}
\footnotesize
\begin{tabularx}{\linewidth}{@{}l r r X@{}}
\toprule
\textbf{Component} & \textbf{Line \%} & \textbf{Branch \%} & \textbf{Notes} \\ \midrule
OCR extractor & 85 & 75 & \\
VAT rules engine & 90 & 85 & Critical path \\
Invoice validator & 90 & 85 & Schema + business rules \\
AP workflow engine & 80 & 70 & \\
\textbf{Overall minimum} & \textbf{85} & \textbf{75} & Enforced in CI \\
\bottomrule
\end{tabularx}
\end{table}

\section{Test fixtures}

Fixtures are located in \texttt{tests/fixtures/arabic/}.

\subsection{Invoice fixtures}

\begin{table}[htbp]
\centering
\caption{Required invoice fixture categories.}
\footnotesize
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Category} & \textbf{Count} & \textbf{Coverage} \\ \midrule
Standard invoices & 20 & 5 per country (SA, EG, BH, QA) \\
Edge cases & 15 & Missing fields, unusual formats \\
Invalid invoices & 10 & Schema violations, invalid TINs \\
Credit notes & 5 & Negative amounts, adjustments \\
Multi-currency & 5 & Non-local currency invoices \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{OCR fixtures}

OCR fixtures cover all confidence thresholds:
\begin{itemize}
  \item High quality ($>$85\%): 10 samples
  \item Medium quality (55--85\%): 10 samples
  \item Low quality (30--55\%): 5 samples
  \item Very low quality ($<$30\%): 5 samples
\end{itemize}

\section{Unit test cases}

\subsection{Invoice extraction}

\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{ID} & \textbf{Priority} & \textbf{Description} \\ \midrule
AP-EXT-001 & P1 & Extract all fields from standard Saudi invoice \\
AP-EXT-002 & P1 & Extract all fields from standard Egyptian invoice \\
AP-EXT-003 & P1 & Extract all fields from standard Bahrain invoice \\
AP-EXT-004 & P1 & Extract all fields from Qatar payment instruction \\
AP-EXT-005 & P1 & Handle missing optional fields gracefully \\
AP-EXT-006 & P2 & Handle multi-page invoice \\
AP-EXT-007 & P2 & Handle poor quality scan ($<$55\% confidence) \\
AP-EXT-008 & P2 & Handle Arabic-only invoice \\
AP-EXT-009 & P2 & Handle Arabic+English bilingual invoice \\
AP-EXT-010 & P3 & Handle unusual date formats \\
\bottomrule
\end{tabularx}

\subsection{VAT validation}

\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{ID} & \textbf{Priority} & \textbf{Description} \\ \midrule
AP-VAT-001 & P1 & Validate 15-digit Saudi TIN format \\
AP-VAT-002 & P1 & Validate 9-digit Egyptian TRN format \\
AP-VAT-003 & P1 & Calculate VAT amount matches invoice \\
AP-VAT-004 & P1 & Apply correct VAT rate by country \\
AP-VAT-005 & P1 & Identify reverse-charge transactions \\
AP-VAT-006 & P2 & Handle zero-rated supplies \\
AP-VAT-007 & P2 & Handle exempt supplies \\
AP-VAT-008 & P2 & Apply withholding tax rules \\
AP-VAT-009 & P3 & Handle credit notes correctly \\
AP-VAT-010 & P3 & Handle inter-GCC transactions \\
\bottomrule
\end{tabularx}

\section{Integration test suite}

\begin{lstlisting}[basicstyle=\ttfamily\small]
pytest tests/integration/arabic/ -v --tb=short
\end{lstlisting}

Expected duration: 15 minutes.

\subsection{End-to-end scenarios}

\begin{enumerate}
  \item \textbf{Happy path}: Invoice scan $\rightarrow$ OCR $\rightarrow$ extraction
        $\rightarrow$ VAT validation $\rightarrow$ approval workflow
  \item \textbf{Manual review}: Low-confidence extraction triggers human review
  \item \textbf{VAT hold}: Rule failure triggers hold and escalation
  \item \textbf{Multi-country}: Invoices from all four countries processed correctly
\end{enumerate}

\section{Performance benchmarks}

All p95 targets in Table~\ref{tab:ap-perf-targets} are measured on the
\emph{reference hardware profile} in Section~\ref{sec:ap-perf-hardware}.
CI enforces each target by failing the performance regression job when
the measured p95 exceeds the target by more than 20\,\% on three
consecutive runs on identical hardware.

\begin{table}[htbp]
\centering
\caption{Arabic AP performance targets on reference hardware.}
\label{tab:ap-perf-targets}
\footnotesize
\begin{tabularx}{\linewidth}{@{}l r r X@{}}
\toprule
\textbf{Operation} & \textbf{p95 Target} & \textbf{Throughput} & \textbf{Notes} \\ \midrule
OCR extraction & 3,000 ms & 200/hr & Per page \\
Field extraction (LLM) & 5,000 ms & 100/hr & Per invoice \\
VAT rule evaluation & 100 ms & 1,000/sec & No external API \\
Full pipeline E2E & 10,000 ms & 50/hr & Including all steps \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Reference hardware profile}
\label{sec:ap-perf-hardware}

\begin{description}
  \item[GPU.] 1$\times$ NVIDIA A100 40\,GB SXM4 \emph{or} L40S 48\,GB.
        Shared with Simula and TB; reservation is not required for
        single-invoice latency targets.
  \item[OCR engine.] Tesseract 5.3.x with Arabic language pack
        (\texttt{ara}, \texttt{ara\_old}), CPU-only. A 2-page invoice
        at 300\,dpi is the reference document size.
  \item[LLM.] \texttt{Qwen2.5-14B-Instruct} served via vLLM 0.6.x,
        \texttt{max\_model\_len=8192}, 14-field extraction prompt
        ($\sim$1,200 input tokens, $\sim$400 output tokens).
  \item[CPU / RAM.] 8~vCPU, 32\,GB RAM host.
  \item[Storage.] NVMe SSD; invoice PDFs streamed from object storage
        with 2\,Gbps link.
  \item[Network.] Co-located vLLM; RTT $<$0.5\,ms. For remote LLM
        endpoints, add measured RTT to the field-extraction p95.
\end{description}

\noindent
Running on consumer-grade GPUs (e.g.\ T4 15\,GB) is not supported for
the 14B model; a fallback profile using
\texttt{Qwen2.5-7B-Instruct} is documented in
\path{docs/arabic/runbooks/perf-fallback.md} with correspondingly
relaxed p95 targets (field extraction $\le$8,000\,ms).

\section{Security tests}

\subsection{Prompt injection}

\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Test} & \textbf{Description} \\ \midrule
AP-SEC-001 & Malicious text in invoice description field \\
AP-SEC-002 & JSON injection via seller name \\
AP-SEC-003 & Unicode control characters in OCR output \\
AP-SEC-004 & System prompt extraction attempt \\
\bottomrule
\end{tabularx}

\section{CI/CD integration}

\begin{lstlisting}[basicstyle=\ttfamily\small,caption={GitHub Actions workflow excerpt.}]
arabic-tests:
  runs-on: ubuntu-latest
  steps:
    - uses: actions/checkout@v4
    - name: Run Arabic AP tests
      run: |
        pytest tests/unit/arabic/ -v --cov=src/arabic
        pytest tests/integration/arabic/ -v
    - name: Check coverage
      run: |
        coverage report --fail-under=85
\end{lstlisting}

\section{Acceptance criteria}

Tests pass when:
\begin{enumerate}
  \item All P1 test cases pass (100\%)
  \item P2 test cases pass ($\geq$95\%)
  \item P3 test cases pass ($\geq$90\%)
  \item Line coverage $\geq$85\%
  \item Performance within targets
  \item No critical security findings
\end{enumerate}