\chapter{Agent Implementation Instructions}\label{ch:agent-impl}

\section{Purpose and audience}
This chapter is the self-contained, intent-based instruction set for
implementing the \emph{Arabic AP invoice processing} domain inside the
unified SAP OSS platform. It is written for LLM-driven and human
implementation agents who will extend existing services to realise the
specification. It assumes only this PDF; every platform decision,
integration contract, and governance obligation the Arabic workstream
depends on is restated here so that no companion document is required.
Existing chapters~\ref{ch:overview}--\ref{ch:ap-roadmap} remain the
normative source for the Arabic domain itself; this chapter tells
future agents \emph{how} to build what those chapters describe,
\emph{where} to put it in the codebase, and \emph{how} it plugs into
the wider platform.

\input{../shared-intent-core.tex}

\section{FMEA: Failure Mode and Effects Analysis for AI}
\label{sec:fmea-ai}

The following analysis identifies critical AI failure modes within the Arabic AP pipeline and their deterministic mitigations.

\begin{table}[htbp]
\centering
\footnotesize
\caption{FMEA for Arabic AP AI Pipeline.}
\label{tab:arabic-fmea}
\begin{tabularx}{\textwidth}{@{}l X X X@{}}
\toprule
\textbf{Failure Mode} & \textbf{Symptom} & \textbf{Impact} & \textbf{Technical Mitigation} \\
\midrule
OCR Character Drift & Misreading '5' as '6' in currency fields. & High (Financial) & Cross-field arithmetic validation (Net + VAT == Gross). \\
Translation Ambiguity & 'Invoice' translated as 'Note' (misclassification). & Medium (Routing) & Metadata-tagging based on ZATCA QR-code extraction. \\
LLM Field Extraction & Extracting 'Address' into 'Vendor Name'. & Medium (Data Qual.) & Confidence-score threshold check per field. \\
\bottomrule
\end{tabularx}
\end{table}

\section{This domain in context}
\label{sec:agent-domain-context}
Arabic AP is the highest-volume transactional workstream among the
four domains. Its primary persona is the AP Clerk, whose working view
is the approval queue surfaced through the shared workspace shell and
gated by AP entitlements. The domain's entry points are document
intake (email, SFTP, upload) producing raw invoice artefacts; its
exit points are (a)~a posted AP document in SAP S/4HANA with a signed
approval record, and (b)~a VAT compliance checklist emitted to the
audit store. Every artefact is validated against the Arabic-owned
schemas in \texttt{structured/schema/} before it leaves the service.
The domain depends on the horizontal regulations wrapper for runtime
governance and on Simula for the training corpus used by the OCR,
field-extraction, and translation models described in
chapter~\ref{ch:ai}.

\section{Schema contribution}
\label{sec:agent-schema-contrib}
Arabic AP owns the schemas catalogued in chapter~\ref{ch:schema}. The
canonical set is the invoice record (see~\ref{tab:invoice-fields}),
the VAT compliance checklist (see~\ref{tab:checklist-fields}), and the
payment instruction record. These schemas live under
\texttt{docs/latex/specs/arabic/} as normative text and under
\texttt{structured/schema/} as the JSON Schema artefacts consumed at
runtime. Registration in the unified registry is additive: each
schema is listed in
\texttt{src/intelligence/ai-core-pal/data\_products/registry.yaml}
with its URI, owning domain (\texttt{arabic}), security class
(\texttt{confidential}), and LLM routing
(\texttt{vllm-only}), following the ODPS~4.1 pattern. Simula consumes
these schemas directly as inputs to its extraction pipeline; it does
not transform them. The \texttt{requirement.schema.json} files owned
by TB and regulations remain distinct from anything Arabic owns and
are not merged with it.

\section{Persona and MCP allow-list}
\label{sec:agent-persona}
\begin{itemize}
  \item \textbf{Persona id:} \texttt{ap-clerk}.
  \item \textbf{Roles:} invoice intake, VAT review, approval
    progression, exception handling.
  \item \textbf{Entitlements:} read/write on the Arabic invoice
    record; read on the VAT rule set; advance transitions on the
    \texttt{po-eproc} and \texttt{direct-ap} state machines defined
    in chapter~\ref{ch:ap}; evidence-write on the control catalogue
    in~\ref{tab:controls}.
  \item \textbf{UI route:} \texttt{/ap/\{tenant\}/queue} on the shared
    workspace shell; surfaced only to callers whose entitlement set
    includes \texttt{ap-clerk}.
  \item \textbf{MCP tool allow-list:}
    \texttt{invoice.intake}, \texttt{invoice.ocr},
    \texttt{invoice.extract-fields},
    \texttt{invoice.translate-ar-en}, \texttt{vat.validate},
    \texttt{vat.calculate}, \texttt{ap.workflow.advance},
    \texttt{ap.workflow.park}. All other MCP tools are denied by
    default per MAS MGF pillar~1.
\end{itemize}

\section{Agent Rule-Pack Binding}
\label{sec:agent-rule-binding}
Before initiating any task in this domain, the agent MUST load the following rule packs in order:
\begin{enumerate}
  \item \filepath{.clinerules} (Global Standards)
  \item \filepath{src/intelligence/.clinerules} (Regulations Envelope)
  \item \filepath{src/arabic/.clinerules} (Arabic Domain Constraints)
\end{enumerate}

\section{Implementation tasks for this domain}
\label{sec:agent-tasks}
The tasks below are intent-based: each task states the outcome a
future agent must achieve and the location in the existing codebase
where the work belongs. None of them contains code; each task binds
to the Arabic chapters that define the required behaviour.

\subsection*{Task A1 --- OCR and translation calibration}
\textbf{Intent.} Bring Arabic OCR and Arabic--English translation up
to the precision/recall targets in chapter~\ref{ch:ocr-pilot}.

\begin{adrbox}{Why Tesseract + Custom Post-Processor?}
We chose Tesseract 5.0 for the baseline due to local-hosting requirements, but added a custom 'Arabic-Context' post-processor to handle the right-to-left layout drift in scanned PDF tables.
\end{adrbox}

\textbf{In-scope.} Running the pilot sample defined
in~\ref{tab:pilot-sample}; publishing per-field thresholds for
accept/review/reject; wiring the thresholds into the pipeline stages
of chapter~\ref{ch:ai}.
\textbf{Out-of-scope.} Choosing a new OCR engine; altering the
canonical invoice schema.
\textbf{Paths to extend.} OCR and extraction code colocated with the
pipeline stages referenced in chapter~\ref{ch:ai}; fixture store
referenced in chapter~\ref{ch:testing}.
\textbf{Definition of done.} Pilot report committed; thresholds
checked in as configuration; the stage-3 field extractor rejects
records below threshold and routes them to human review.

\textbf{Autonomous Verification (Agent Logic):}
\begin{lstlisting}[language=jsonx]
// Verify Task A1
{
  "test_fixture": "tests/fixtures/arabic/invoice_sample_01.png",
  "expected_outcome": "ocr_confidence > 0.88",
  "validation_command": "make test-arabic-ocr"
}
\end{lstlisting}

\textbf{Verification.} \texttt{make spec-arabic} builds cleanly and
the Arabic test suite described in chapter~\ref{ch:testing} passes
end-to-end on the calibration fixtures.

\subsection*{Task A2 --- Multi-country VAT engine}
\textbf{Intent.} Implement the VAT compliance engine specified in
chapter~\ref{ch:vat} and detailed in chapter~\ref{ch:vat-impl}.
\textbf{In-scope.} Per-country rule modules for Saudi Arabia (ZATCA),
UAE, Bahrain (NBR), and Oman; registration validation, calculation,
activity lookup, and checklist emission; the canonical decision
function in~\ref{fig:ksa-decision}.
\textbf{Out-of-scope.} VAT rules for countries outside the in-scope
list in~\ref{tab:countries}.
\textbf{Paths to extend.} The VAT module layout referenced
in~\ref{fig:vat-arch}; API contracts in chapter~\ref{ch:vat-impl}.
\textbf{Definition of done.} Every rule cited
in~\ref{tab:ksa-rules} has a callable check; the checklist record is
emitted per invoice and validates against
\texttt{vat-checklist.schema.json}.
\textbf{Verification.} Unit tests for each country module (see
chapter~\ref{ch:testing}) pass; the VAT calculation API contract
sample in chapter~\ref{ch:vat-impl} returns the expected response.

\subsection*{Task A3 --- AP approval workflow}
\textbf{Intent.} Realise the AP workflow and approval engine in
chapter~\ref{ch:ap}, including the \texttt{po-eproc} and
\texttt{direct-ap} state machines.
\textbf{In-scope.} Workflow-selection logic in~\ref{fig:workflow-select};
state transitions in~\ref{fig:po-eproc}; park, reroute, and escalate
actions with evidence records.
\textbf{Out-of-scope.} SAP S/4HANA AP posting internals;
non-Bahrain delegations.
\textbf{Paths to extend.} Workflow engine hosted in the gateway
service; evidence store colocated with the control catalogue.
\textbf{Definition of done.} Each state in the state-machine diagram
is reachable from a deterministic predecessor; transitions are audit
logged with the persona id of the acting user; the golden regression
set in chapter~\ref{ch:controls} executes without human intervention.
\textbf{Verification.} Workflow unit tests pass; the approval queue
UI surfaces the correct entitlements for \texttt{ap-clerk}.

\subsection*{Task A4 --- Bahrain DOI seven controls}
\textbf{Intent.} Evidence the seven Bahrain Delegation of Invoicing
controls required in chapter~\ref{ch:controls}.
\textbf{In-scope.} Control enforcement map in~\ref{fig:controls-map};
the seven rows of the control catalogue in~\ref{tab:controls} that
concern the Bahrain DOI; evidence emission per control.
\textbf{Out-of-scope.} Introducing controls for other delegations.
\textbf{Paths to extend.} Control catalogue file referenced in
chapter~\ref{ch:controls}; evidence store.
\textbf{Definition of done.} Running the workflow on the Bahrain
golden sample triggers every control; each control writes an
evidence record conforming to the Arabic schema set.
\textbf{Verification.} Traceability matrix in chapter~\ref{ch:controls}
is complete; the testing chapter's acceptance suite passes.

\subsection*{Task A5 --- Testing and golden regression set}
\textbf{Intent.} Deliver the testing specification in
chapter~\ref{ch:testing} at the coverage targets
in~\ref{tab:arabic-coverage}.
\textbf{In-scope.} Invoice and OCR fixtures; unit, integration, and
regression suites; CI wiring.
\textbf{Out-of-scope.} Performance benchmarking infrastructure beyond
what chapter~\ref{ch:testing} defines.
\textbf{Paths to extend.} Test directory colocated with the Arabic
service code; fixture store used by Simula as training input.
\textbf{Definition of done.} Coverage meets the thresholds
in~\ref{tab:arabic-coverage}; CI gates block regressions; the
regression golden set is reproducible from fixture seeds.
\textbf{Verification.} \texttt{make spec-arabic} remains green; the
Arabic test entry points listed in chapter~\ref{ch:testing} all run.


\section{Governance obligations applied to this domain}
\label{sec:agent-governance}
The regulations domain wraps Arabic AP at runtime. Four obligations
apply without exception; each is stated as a concrete hook in the
Arabic codebase.

\begin{itemize}
  \item \textbf{MAS MGF four pillars} (see regulations
    chapter~3 in \texttt{docs/latex/specs/regulations/chapters/03-mgf-framework.tex}).
    Pillar~1 (assess and bound): every Arabic MCP tool is declared in
    the allow-list in \S\ref{sec:agent-persona}, deny-by-default.
    Pillar~2 (design and deploy): the pipeline stages in
    chapter~\ref{ch:ai} are traceable end-to-end via the
    per-request identity attribution described below. Pillar~3
    (operate and monitor) and pillar~4 (govern and disclose) are
    satisfied by emitting the evidence records produced by the
    Arabic control catalogue (chapter~\ref{ch:controls}) into the
    shared governance store.
  \item \textbf{Identity attribution} (regulations chapter~10,
    \texttt{10-identity-attribution.tex}). Every MCP call and every
    workflow transition carries the persona id, agent id, request
    id, and tenant id. The gateway is the attachment point; Arabic
    code propagates the identity envelope without mutating it.
  \item \textbf{Emergent-capability monitoring} (regulations
    chapter~8, \texttt{08-monitoring-spec.tex}). OCR, extraction,
    translation, and VAT calculation are instrumented with the
    monitoring metrics specified in that chapter; metrics flow to
    the monitoring sink defined there.
  \item \textbf{Conformance tooling} (regulations chapter~6,
    \texttt{06-conformance-tooling.tex}). The Arabic service exposes
    AI Verify 2.0 plug-in hooks and Moonshot CI/CD 1.1.0 benchmark
    entry points as required by that chapter.
\end{itemize}

\section{Cross-spec integration --- full detail}
\label{sec:agent-integration}

\subsection{With Arabic AP}
Arabic AP exports the invoice record, VAT compliance checklist, and
payment instruction schemas under \texttt{structured/schema/}
(chapter~\ref{ch:schema}); registers the MCP tools enumerated in
\S\ref{sec:agent-persona}; and applies the four governance hooks in
\S\ref{sec:agent-governance} to every call it serves. External
consumers (TB, regulations, Simula) see Arabic only through these
contracts; there are no private back-channels.

\subsection{With Trial Balance}
Arabic AP produces posted AP records and VAT checklists that feed
the TB close. TB consumes them as evidence inputs to its
commentary engine, not as schema merges: the Arabic invoice schema
and the TB control record remain distinct URIs in the unified
registry. Integration point: TB's ingestion stage calls the
Arabic MCP tool \texttt{invoice.extract-fields} with read-only
entitlements; the response is validated against
\texttt{invoice.schema.json} before TB accepts it.

\subsection{With Regulations}
Regulations wraps Arabic at runtime. Every Arabic MCP call is
brokered through the regulations governance gate, which applies
pillars~1--4 of MAS MGF, attaches the identity envelope, and routes
emergent-capability metrics to the monitoring store. The Arabic
service code does not call regulations directly; the gateway layer
enforces this. Conformance tests defined in regulations
chapter~5 (empirical evidence) and chapter~6 (conformance tooling)
execute against the Arabic service as part of the platform CI gate.

\subsection{With Simula}
Simula consumes Arabic's schemas as training input. The contract is
schema-in, \texttt{TrainingExample}-out: Simula reads
\texttt{invoice.schema.json}, \texttt{vat-checklist.schema.json}, and
the Arabic fixture store; it produces \texttt{TrainingExample}
records. Simula does not call Arabic MCP tools at runtime and does
not share state with the Arabic workflow. No cross-domain
operational envelope is introduced.

\section{Wave assignment and sequencing}
\label{sec:agent-waves}
Implementation runs in three waves.

\begin{itemize}
  \item \textbf{Wave 1 --- foundations.} Unified schema registry
    entry; persona-gated surface for the AP Clerk; identity
    attribution envelope flowing through the gateway. Prerequisites:
    the regulations chapter~10 identity schema must be frozen.
  \item \textbf{Wave 2 --- Arabic core build-out.} Tasks A1--A4 above.
    Prerequisites: Wave~1 complete; Simula training fixtures
    available from Simula Wave~1 (extraction pipeline).
  \item \textbf{Wave 3 --- hardening and conformance.} Task A5
    and full conformance-tooling hook-up. Prerequisites: Wave~2
    complete; regulations Wave~2 conformance tooling integrated;
    monitoring sink live.
\end{itemize}

Arabic occupies Wave~2 for its core build-out and Wave~3 for its
testing and conformance surfaces.

\section{Acceptance criteria (platform-wide)}
\label{sec:agent-acceptance}
When all four chapter~12 instruction sets are implemented, the
platform satisfies the following invariants:

\begin{enumerate}
  \item \textbf{One instance.} A single codebase runs a single
    logical instance; no domain ships as a separate product.
  \item \textbf{Persona-gated surfaces.} The four personas each
    reach a tailored UI on the shared shell; no route, MCP tool, or
    data product is visible outside its persona's entitlement set.
  \item \textbf{Unified registry.} Every schema in use is listed in
    the ODPS~4.1 registry under its owning domain; validators
    resolve \texttt{\$ref}s through the unified registry only.
  \item \textbf{No schema merges.} The two
    \texttt{requirement.schema.json} files in TB and regulations
    remain distinct; no cross-domain composite schemas are produced.
  \item \textbf{No cross-domain operational envelope.} Runtime
    orchestration stays within each domain's service; only Simula
    consumes other domains' outputs, and only as training input.
  \item \textbf{Governance wraps all three non-regulations domains.}
    Arabic, TB, and Simula service calls pass through the
    regulations runtime wrapper without exception.
\end{enumerate}

\section{References}
\label{sec:agent-references}
Within-spec references:
chapter~\ref{ch:overview} (overview),
chapter~\ref{ch:schema} (data schema),
chapter~\ref{ch:ai} (AI pipeline),
chapter~\ref{ch:vat} (VAT compliance engine),
chapter~\ref{ch:ap} (AP workflow),
chapter~\ref{ch:controls} (controls and testing),
chapter~\ref{ch:references} (references),
chapter~\ref{ch:vat-impl} (VAT implementation),
chapter~\ref{ch:testing} (testing specification),
chapter~\ref{ch:ocr-pilot} (OCR pilot),
chapter~\ref{ch:ap-roadmap} (project roadmap).

Peer chapter~12 instruction sets for the other three domains (each
is self-contained; consult directly when operating in that domain):
\begin{itemize}
  \item Trial Balance:
    \texttt{docs/latex/specs/tb/chapters/12-implementation-instructions.tex}.
  \item Regulations:
    \texttt{docs/latex/specs/regulations/chapters/12-implementation-instructions.tex}.
  \item Simula:
    \texttt{docs/latex/specs/simula/chapters/12-implementation-instructions.tex}.
\end{itemize}

Shared platform artefacts:
\begin{itemize}
  \item ODPS~4.1 registry:
    \texttt{src/intelligence/ai-core-pal/data\_products/registry.yaml}.
  \item Schema pipeline and resolver:
    \texttt{src/intelligence/schema\_pipeline/}.
  \item Training pipeline (Simula):
    \texttt{src/training/pipeline/}.
  \item Shared workspace shell:
    \texttt{generativeUI/training-webcomponents-ngx}.
\end{itemize}
