\chapter{References}\label{ch:references}

This appendix lists every file that backs a normative statement in the
specification. Paths are relative to the repository root. SHA-256
hashes are taken from \corpus\texttt{structured/corpus.json} and are
the ground truth for the provenance check of \cref{ch:schema}.

\section{Source documents}

\footnotesize
\begin{longtable}{@{}p{3.4cm}p{8cm}p{3cm}@{}}
\toprule
\textbf{Short name} & \textbf{Path} & \textbf{SHA-256 (prefix)} \\ \midrule
\endfirsthead
\toprule
\textbf{Short name} & \textbf{Path} & \textbf{SHA-256 (prefix)} \\ \midrule
\endhead
TB GLO control narrative & \path{docs/tb/source/Operation of Month End Close Controls  - Trial Balance Review - GLO (1).pdf} & \texttt{71a4524b9ec7\ldots} \\
TB Business Case         & \path{docs/tb/source/Business case - Trial Balance (2).doc} & \texttt{32744df8ea13\ldots} \\
BSS RA Business Case     & \path{docs/tb/source/Business case - BSS Risk Assessment (2).doc} & \texttt{dab433cecf59\ldots} \\
DOI for TB Process       & \path{docs/tb/source/DOI for Trial Balance Process .docx} & \texttt{b9d52d365b6b\ldots} \\
HKG TB Workbook (Nov'25) & \path{docs/tb/source/HKG_TB review Nov'25.xlsm} & \texttt{69f5da1f4c71\ldots} \\
HKG PL Workbook (Nov'25) & \path{docs/tb/source/HKG_PL review Nov'25.xlsm} & \texttt{47716e8c107b\ldots} \\
\bottomrule
\end{longtable}
\normalsize

\section{Derived text}

Extracted text for each source document lives at
\corpus\texttt{extracted-text/}. The Word and PDF files are converted
to Markdown; the macro-enabled workbooks produce CSV samples (one per
book) that are sufficient to seed downstream joins without carrying
the macro payload. Extraction uses \texttt{pandoc} for \texttt{.docx},
\texttt{pdftotext -layout} for born-digital PDFs, and a small Python
pipeline for the \texttt{.xlsm} sheets.

\section{Structured records}

All of the following are JSON records validated by
\corpus\texttt{structured/validate.py}.

\footnotesize
\begin{longtable}{@{}p{4cm}p{10cm}@{}}
\toprule
\textbf{Type} & \textbf{Path} \\ \midrule
\endhead
Workflow          & \path{structured/workflows/tb-review-glo.json} \\
Control           & \path{structured/controls/euc-001-tb-data-extraction.json} \\
Control           & \path{structured/controls/euc-002-variance-analysis-file.json} \\
Control           & \path{structured/controls/euc-003-internal-review.json} \\
Decision point    & \path{structured/decision-points/dp-001-material-variance.json} \\
Decision point    & \path{structured/decision-points/dp-002-journal-posting.json} \\
Decision point    & \path{structured/decision-points/dp-003-further-investigation.json} \\
Decision point    & \path{structured/decision-points/dp-004-entry-confirmation.json} \\
Decision point    & \path{structured/decision-points/dp-005-unexplained-variance.json} \\
Requirement       & \path{structured/requirements/tb-business-requirements.json} \\
KPI collection    & \path{structured/kpis/tb-kpis.json} \\
Glossary          & \path{structured/glossary/tb-glossary.json} \\
Data product      & \path{structured/data-products/trial-balance-review-v1.json} \\
Index             & \path{structured/corpus.json} \\
Enumerations      & \path{structured/enums.yaml} \\
\bottomrule
\end{longtable}
\normalsize

\section{Tools}

\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Tool} & \textbf{Purpose} \\ \midrule
\texttt{pandoc}   & \texttt{.docx} $\to$ Markdown. \\
\texttt{pdftotext} -layout & Text extraction from the GLO control PDF. \\
Python + \texttt{openpyxl} & Sample rows out of \texttt{.xlsm} workbooks. \\
\texttt{jq} & Query and cross-check \texttt{corpus.json}. \\
\texttt{jsonschema} / \texttt{referencing} (Python) & Offline schema validation with a local registry. \\
\texttt{xelatex} / \texttt{tectonic} & Build this document. \\
\bottomrule
\end{tabularx}

\section{Minimal corpus excerpt}

The entire corpus is in-tree under \corpus. For readers without access
to the Git repository, \cref{lst:corpus-excerpt} reproduces the first
three items of \corpus\texttt{structured/corpus.json} verbatim.

\begin{figure}[htbp]
\begin{lstlisting}[language=jsonx,caption={First three items of
\texttt{corpus.json}; identical SHA-256 values are reproduced in
the table above.},label=lst:corpus-excerpt]
{
  "schema_version": "1.0",
  "items": [
    {
      "id": "tb-glo-control-operation",
      "type": "control_narrative",
      "country": "GLO",
      "language": "en",
      "sha256": "71a4524b9ec7c6701a7137579ef6c8730eb53fcde2fe5a50515a4b870b560afc",
      "source_path":     "docs/tb/source/Operation of Month End Close Controls ...GLO (1).pdf",
      "text_path":       "docs/tb/extracted-text/bpmn-tb-review-glo.md",
      "structured_path": "docs/tb/structured/workflows/tb-review-glo.json"
    },
    {
      "id": "tb-business-case",
      "type": "business_case",
      "country": "GLO",
      "language": "en",
      "sha256": "32744df8ea13f12a36f6117f18874e578c627be6fbc3c48d2f73d32dcfa0c689",
      "source_path":     "docs/tb/source/Business case - Trial Balance (2).doc",
      "text_path":       "docs/tb/extracted-text/business-case-trial-balance.md",
      "structured_path": "docs/tb/structured/requirements/tb-business-requirements.json"
    },
    {
      "id": "bss-ra-business-case",
      "type": "business_case",
      "country": "GLO",
      "language": "en",
      "sha256": "dab433cecf598b908d070e2ebfd6f285565b7b3e1102cd2496e639dee1d0c28c",
      "source_path": "docs/tb/source/Business case - BSS Risk Assessment (2).doc"
    }
  ]
}
\end{lstlisting}
\end{figure}

\section{Out-of-scope assets}

The BSS Risk Assessment business case
(\path{docs/tb/source/Business case - BSS Risk Assessment (2).doc})
is kept in the corpus for terminology alignment but is out of scope
for this specification; it is covered by a sibling business case with
its own lifecycle.
