% =============================================================================
% Frontmatter --- abstract, document purpose, audience
% =============================================================================

\chapter*{Document status and purpose}
\addcontentsline{toc}{chapter}{Document status and purpose}

This specification formalises the regulations, frameworks, and conformance
tooling relevant to agentic AI deployments in the SAP OSS finance stack. Its
three goals are:

\begin{enumerate}
  \item \textbf{Make the regulations machine-readable.} Every source document
        in \corpus{source/} is accompanied by structured JSON records under
        \corpus{structured/} that capture sections, normative requirements,
        and observational claims.
  \item \textbf{Bind each requirement to evidence in the repository.} The
        requirement schema carries an \texttt{implementation[]} array of file
        paths, configuration fields, and documentation anchors, plus a
        \texttt{verification[]} array pointing to AI Verify plugins, Moonshot
        benchmark IDs, or repository tests. A local validator
        (\corpus{structured/validate.py}) fails when any of these paths no
        longer resolve.
  \item \textbf{Give reviewers a single printable reference.} This document
        cross-links the corpus records, the two vendored conformance
        frameworks, and the code they exercise, so that governance reviews,
        internal audit, and external assessors can read one PDF instead of
        navigating three directory trees.
\end{enumerate}

\paragraph{In scope.} The Monetary Authority of Singapore (MAS) / AI Verify
Foundation \emph{Model AI Governance Framework for Agentic AI} v1.0 (22 Jan
2026); the multi-institution \emph{2025 AI Agent Index}; the Ju \& Aral field
experiment \emph{Collaborating with AI Agents} (arXiv 2503.18238v3, Feb 2026);
and the two vendored open-source conformance tools
\tool{aiverify} (v2.0) and \tool{moonshot-cicd} (v1.1.0).

\paragraph{Out of scope.} Binding legislation (EU AI Act, US executive
orders); sector-specific prudential rules (Basel, Solvency); deterministic
non-LLM agent platforms. Those are tracked separately and may be added as
additional regulation records without changing the schema.

\paragraph{Audience.} AI governance leads, internal audit, model risk, and
technical reviewers responsible for evidencing conformance for agentic
deployments that use the SAP AI Core / PAL / BTP stack captured in
\corpus{../tb/} and \corpus{../arabic/}.

\paragraph{Relationship to other corpora.} This corpus is a peer of
\corpus{../arabic/} (Arabic invoice processing) and \corpus{../tb/} (trial
balance review). Where requirements apply to those workloads, the
\texttt{implementation[]} field points back to their structured records or
spec chapters.

\vspace{1em}
\noindent\textcolor{reggrey}{\small This is a living document. Re-running
\tool{python3 docs/regulations/structured/validate.py} checks schema
conformance and verifies that every code path, plugin, and test id
referenced from the requirement records still exists.}
\clearpage
