\chapter{Source Analysis: Bahrain AP DOI}
\label{ch:analysis-bh-doi}

\section{Document Context}
The \texttt{Accounts Payable DOI.docx} (Bahrain) defines the delegation of authority and mandatory controls for invoice processing. This document was processed into \texttt{structured/policies/bh-ap-doi.json}.

\section{Detailed Requirement Mapping}

\begin{longtable}{p{5cm} p{5cm} p{5cm}}
\caption{Bahrain DOI: Source-to-Requirement Lineage} \\
\toprule
\textbf{Source Paragraph} & \textbf{Business Requirement} & \textbf{Formal Requirement (Spec)} \\
\midrule
\endfirsthead
\toprule
\textbf{Source Paragraph} & \textbf{Business Requirement} & \textbf{Formal Requirement (Spec)} \\
\midrule
\endhead
Section 4.1: ``All invoices must be matched against a valid Purchase Order prior to processing.'' & \textbf{BR-DOI-M1:} Mandatory PO-Matching for standard AP flow. & \texttt{arabic-ap-spec} Ch. 5 (PO-EPROC state machine). \\
\midrule
Section 4.2: ``Any variance > 5\% between PO and Invoice must be routed to the Finance Head.'' & \textbf{BR-DOI-V1:} Automated variance threshold detection. & \texttt{arabic-ap-spec} Ch. 6 (Variance Control C3). \\
\midrule
Section 4.7: ``The final payment file must be hashed and signed by two distinct personas.'' & \textbf{BR-DOI-S1:} Dual-Persona Cryptographic Sign-off. & \texttt{arabic-ap-spec} Ch. 12 (HITL Gate: Payment Instruction). \\
\bottomrule
\end{longtable}

\section{Machine-Readable Evidence (Fragment)}
The following YAML fragment was extracted from the DOI to drive the 28-state workflow engine:

\begin{lstlisting}[language=jsonx]
{
  "policy_id": "BH-DOI-2026",
  "controls": [
    {
      "id": "C3",
      "description": "PO-Invoice Variance check",
      "threshold": 0.05,
      "escalation_persona": "finance_head"
    }
  ]
}
\end{lstlisting}
