\chapter{Overview and scope}\label{ch:overview}

\section{The problem}

Standard Chartered Bank operates Accounts Payable across a heterogeneous
regulatory and linguistic landscape. Depending on the country, an
incoming invoice may be:

\begin{itemize}
  \item issued by a national e-invoicing portal (Egypt's ETA, with a
        26-character \textsc{electronic\_id}; Saudi Arabia's ZATCA with
        an analogous assessment reference);
  \item purely Arabic (as in the two Egyptian invoices
        \texttt{3CPJR5CQ\ldots K10.pdf} and \texttt{B1STE8RH\ldots SJ10.pdf}
        and the Saudi VAT assessment \ar{اشعار صدور فاتورة});
  \item bilingual (Qatar's QCSD / EDAA memo);
  \item wholly English with pre-assigned GL coding (e.g.\ the Bahrain
        DOI-covered PO flow).
\end{itemize}

Today the Arabic-language material is handled manually: a finance analyst
reads the PDF, transcribes key fields into an OTP/VAT checklist (see the
Saudi Q3 2025 checklist in \corpus\texttt{source/Saudi\ldots OTP\ldots}),
and routes it through the DOI approval chain. The November 2025 e-mail
thread from Manoj Hemath Kumar identifies this manual path as the target
for AI replacement.

\section{Goals}

The specification targets three outcomes, each with a direct lineage to
the sample corpus in \corpus\texttt{source/}:

\begin{enumerate}
  \item \textbf{Structured ingestion.} Every Arabic or bilingual invoice
        becomes a record conforming to
        \texttt{structured/schema/invoice.schema.json}
        (\cref{ch:schema}).
  \item \textbf{Deterministic compliance.} Every invoice is passed
        through a rule engine that emits a VAT compliance checklist
        conforming to \texttt{structured/schema/vat-checklist.schema.json}
        (\cref{ch:vat}), with the rule set derived from the
        regulator-specific controls in the corpus.
  \item \textbf{Auditable approval.} Every payment follows one of the
        four state machines in \corpus\texttt{structured/workflows/} and
        meets the controls enumerated in the Bahrain DOI
        (\cref{ch:ap,ch:controls}).
\end{enumerate}

\section{In-scope countries and counterparties}

\begin{table}[htbp]
\centering
\caption{Countries in scope, with the controlling source document and
         the structured records derived from it.}
\label{tab:countries}
\small
\begin{tabularx}{\linewidth}{@{}l l l X@{}}
\toprule
\textbf{Country} & \textbf{Authority} & \textbf{Currency} &
\textbf{Source document(s)} \\ \midrule
Egypt (EG)  & ETA   & EGP &
  \texttt{3CPJR5CQ\ldots K10.pdf},\newline\texttt{B1STE8RH\ldots SJ10.pdf} \\
Saudi Arabia (SA) & ZATCA & SAR &
  \ar{اشعار صدور فاتورة} (005).pdf, \newline Saudi OTP/VAT Q3'25 checklist \\
Qatar (QA)  & QTA   & QAR &
  EDAA Annual Fee 2025, \newline QCSD payment instruction \\
Bahrain (BH) & NBR  & BHD &
  Accounts Payable DOI \\
Iraq (IQ), UAE, PK & various & --- &
  \emph{Planned}; referenced in the initiative e-mail but formally
  out of scope until at least five sample invoices per jurisdiction
  are added to \texttt{source/} \\ \bottomrule
\end{tabularx}
\end{table}

\section{System context}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  node distance=9mm and 13mm,
  box/.style={draw,rounded corners,minimum width=26mm,minimum height=9mm,
              align=center,font=\small},
  ext/.style={box,fill=blue!6},
  sys/.style={box,fill=green!7},
  ai/.style={box,fill=orange!10},
  ap/.style={box,fill=violet!8},
  arr/.style={-{Latex[length=2.2mm]},thick}
]
\node[ext] (eta)   {ETA portal\\(Egypt)};
\node[ext,right=of eta] (zatca) {ZATCA portal\\(Saudi Arabia)};
\node[ext,right=of zatca] (vendor) {Vendor PDFs\\e-mail/scan};

\node[ai,below=12mm of zatca] (ai) {Arabic-invoice\\AI pipeline\\(Ch.~\ref{ch:ai})};

\node[sys,below=of ai,xshift=-28mm] (vat) {VAT engine\\(Ch.~\ref{ch:vat})};
\node[sys,below=of ai,xshift= 28mm] (wf)  {Approval engine\\(Ch.~\ref{ch:ap})};

\node[ap,below=14mm of vat,xshift=28mm] (gl) {GL posting\\(out of scope)};

\draw[arr] (eta)    -- (ai);
\draw[arr] (zatca)  -- (ai);
\draw[arr] (vendor) -- (ai);
\draw[arr] (ai)     -- (vat);
\draw[arr] (ai)     -- (wf);
\draw[arr] (vat)    -- (gl);
\draw[arr] (wf)     -- (gl);
\end{tikzpicture}
\caption{System context. Regulator portals (blue) and vendor PDFs feed
         the AI pipeline (orange), which emits canonical records to
         the deterministic VAT and approval engines (green). Downstream
         posting to GL is out of scope.}
\label{fig:context}
\end{figure}

\section{Out of scope}

The following are \emph{explicitly} out of scope for this document:

\begin{itemize}
  \item Integration contracts with PeopleSoft PSAP/PSGL, eBBS, eOPS, and
        Tax-Q. These are covered by a separate interface specification.
  \item Vendor master-data on-boarding and the 48-hour screening service
        itself; this document assumes a screening result is available as
        a boolean precondition.
  \item The encrypted template
        \texttt{Business Instruction Template - V1.0 (002) (1).xlsm} and the
        duplicate \texttt{EDAA\ldots[65].pdf}; neither is referenced by
        any structured record.
  \item Peer/competitor analysis across jurisdictions, which the e-mail
        thread itself identifies as a separate initiative.
\end{itemize}

\section{Five-session engagement model}

Per the e-mail, every country instance of this system follows a
five-session agile discovery cadence
(\Cref{tab:sessions}); the AI pipeline of \cref{ch:ai} is designed to
advance with that cadence.

\begin{table}[htbp]
\centering
\caption{Five-session engagement model (codified in
         \corpus\texttt{structured/enums.yaml}).}
\label{tab:sessions}
\small
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Code} & \textbf{Name} & \textbf{Goal} \\ \midrule
\textsc{understand} & Session~1 & Map problem, define user journey, pick target. \\
\textsc{sketch}     & Session~2 & Develop detailed solution design. \\
\textsc{decide}     & Session~3 & Critique approaches, select best path. \\
\textsc{prototype}  & Session~4 & Build practical system for use case. \\
\textsc{test}       & Session~5 & Test with $\geq$5 stakeholders, gather feedback. \\
\bottomrule
\end{tabularx}
\end{table}
