\documentclass{sap-oss-apple-minimal}

\title{Test Document}
\author{SCB GCFO, AI Research \& Development}
\date{April 2026}
\setdocsubtitle{WWDC Quality Style Demonstration}
\setdocref{TEST-001}
\setdocversion{3.0.0}
\setbreadcrumbroot{SAP OSS Specs}

\begin{document}

\makesaptitle

\tableofcontents
\cleardoubleemptypage

\chapter{Introduction}

This document demonstrates the WWDC-quality styling for SAP-OSS technical specifications. The design prioritizes readability, visual hierarchy, and professional presentation.

\section{Design Philosophy}

The styling follows Apple's WWDC keynote aesthetic:
\begin{itemize}
  \item Clean typography with generous whitespace
  \item Dramatic chapter openers with large faded numbers
  \item Breadcrumb navigation in headers
  \item Subtle shadows and accent colors
  \item Minimal tables with thin rules
\end{itemize}

\subsection{Typography}

Body text uses Inter (or Helvetica as fallback) at 10pt with 1.35× line height and a full-baseline paragraph skip. This provides excellent readability with generous breathing room between paragraphs.

\subsubsection{Heading Hierarchy}

Headings use a clear size progression: 28pt for chapters, 18pt for sections, 14pt for subsections, and 12pt for subsubsections.

\section{Box Environments}

\begin{adrbox}[ADR-001: Use XeLaTeX for Compilation]
All documents must be compiled with XeLaTeX to ensure proper font rendering and Unicode support. PDFLaTeX is not supported.
\end{adrbox}

\begin{infobox}[Key Information]
The document class automatically loads common packages including TikZ, booktabs, tcolorbox, and fontawesome5.
\end{infobox}

\begin{warningbox}[Caution]
Ensure the Inter and JetBrains Mono fonts are installed on your system for optimal rendering.
\end{warningbox}

\begin{successbox}[Validation Passed]
All tests completed successfully. The document class is ready for production use.
\end{successbox}

\begin{criticalbox}[Breaking Change]
This version (v3) is not backward compatible with v1. Update your document preambles accordingly.
\end{criticalbox}

\begin{notebox}[Implementation Note]
The chapter opener uses TikZ overlays to position the large faded chapter number.
\end{notebox}

\section{Status Badges}

Requirements can be tagged with MoSCoW priorities. The \texttt{badgelist} environment keeps descriptors aligned in a consistent column regardless of badge width:

\begin{badgelist}
  \badgeitem{\statusmust}{Non-negotiable for v1.0. Ship is blocked without these.}
  \badgeitem{\statusshould}{Important but can slip to a later release if needed.}
  \badgeitem{\statuscould}{Nice to have once the core is stable and measured.}
  \badgeitem{\statuswont}{Explicitly out of scope for this phase.}
\end{badgelist}

\section{Helper Commands}

\begin{itemize}
  \item File paths: \filepath{src/intelligence/analytics/}
  \item Artifacts: \artifact{.clinerules}
  \item Requirement IDs: \reqid{TB-REQ-ES01}
  \item Schema paths: \schemapath{docs/schema/tb/variance-record.schema.json}
  \item Enum values: \enumval{approved}
  \item Field references: \fieldref{variance\_id}
  \item Tools: \tool{xelatex}
  \item States: \stateid{COMPLETED}
\end{itemize}

\subsection{SF Symbols 7}

The helper \texttt{\textbackslash sfsymbol\{<hex>\}} drops any SF Symbols~7 glyph inline using the installed \emph{SF Pro} font. Codepoints are found in the \emph{SF Symbols} app via \emph{Edit \textbar{} Copy \textbar{} Copy as Unicode}, then pasted without the leading \texttt{0x}:

\begin{badgelist}
  \badgeitem{\sfsymbol{100181}}{Whatever glyph your SF Symbols 7 install maps to \texttt{0x100181}.}
  \badgeitem{\sfsymbol{100195}}{Glyph at \texttt{0x100195} --- codepoints shift per SF Symbols release.}
  \badgeitem{\sfsymbol{1001F8}}{Glyph at \texttt{0x1001F8}.}
  \badgeitem{\sfsymbol{100184}}{Glyph at \texttt{0x100184}.}
\end{badgelist}

The helper is agnostic to version; copy any codepoint from your SF Symbols app and it will render.

\chapter{Tables and Code}

\section{Metrics Table}

\begin{metricstable}{Key Performance Metrics}
Accept without edit rate & 72\% & $\geq$70\% \\
Accept with any edit rate & 96\% & $\geq$95\% \\
vLLM availability & 99.2\% & $\geq$95\% \\
vLLM latency P95 & 3200ms & $<$5000ms \\
\end{metricstable}

\section{Code Listings}

Source code blocks render with an Xcode-style file-path header, line numbers, and syntax highlighting:

\begin{codeblock}[language=Python]{src/training/pipeline/simula_config.py}
from dataclasses import dataclass
from pathlib import Path

@dataclass(frozen=True)
class SimulaConfig:
    """Configuration for the Simula training pipeline."""
    corpus_path: Path
    batch_size: int = 32
    learning_rate: float = 3e-5

    def validate(self) -> None:
        if not self.corpus_path.exists():
            raise FileNotFoundError(self.corpus_path)
\end{codeblock}

\section{Source Quotations}

\begin{sourcequote}
``The Trial Balance Review process ensures that account balances are accurate, supported, and reconciled before period close.'' --- TB-BC-001, Section 2.1
\end{sourcequote}

\makecolophon{%
This document was typeset with XeLaTeX using the sap-oss-apple-minimal document class.\\[8pt]
Version 3.0.0 --- April 2026\\[8pt]
SCB GCFO, AI Research \& Development
}

\end{document}